---
title: "10 Tibbles"
author: "Erika Anderson"
date: "May 8, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load library
library(tidyverse)

```

A group of DFO coworkers decided to work together to compete the exercises in [R for Data Science](https://r4ds.had.co.nz/) by Garrett Grolemund and Hadley Wickham. Many of the solutions to the exercises are available online, with the most [extensive version](https://jrnold.github.io/r4ds-exercise-solutions/index.html) that I have found so far by Jeffrey Arnold on Git Hub.

# 10.5 Exercises

1. How can you tell if an object is a tibble? (Hint: try printing mtcars, which is a regular data frame).

```{r}

# print original data
print(mtcars)

# check if tibble
is_tibble(mtcars)

# convert into tibble
mtcarstib <- as_tibble(mtcars)

# check if tibble
is_tibble(mtcarstib)

# use class function
class(mtcars)
class(mtcarstib)

# print tibble
print(mtcarstib)
```


2. Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?

- data frames return imperfect matching column name
- subsets returns with column name
- data types are given in tibbles results

```{r}

df <- data.frame(abc = 1, xyz = "a")
dftib <- as_tibble(df)

df$x # incorrectly returns column xyz
dftib$x # gives error

df[, "xyz"] # returns a
dftib[, "xyz"] # returns column xyz with a

df[, c("abc", "xyz")] # returns columns
dftib[, c("abc", "xyz")] # returns columns and data type

```


3. If you have the name of a variable stored in an object, e.g. var <- "mpg", how can you extract the reference variable from a tibble?

```{r}

var <- "mpg"

mtcarstib <- as_tibble(mtcars)

# subset
mtcarstib[var]

# another way
mtcarstib %>%
  select(var)

# these do not work with a column name stored in a variable
# # using dollar sign
#carstib$var
# 
# # with pipe
# carstib %>%
#   .$var

```


4. Practice referring to non-syntactic names in the following data frame by:

```{r}

annoying <- tibble(
  `1` = 1:10,
  `2` = `1` * 2 + rnorm(length(`1`))
)

```
    
  + Extracting the variable called 1.
  
```{r}

# base R way to pull observation one
annoying[[1,2]]

# another way to pull column one
annoying$`1`

# another way to pull column one
annoying %>%
filter(., `1` == 1) %>%
  pull(`2`)

```
  

  + Plotting a scatterplot of 1 vs 2.
  
```{r}
ggplot(annoying, aes(`1`, `2`)) +
  geom_point()

# what happens with no back ticks
ggplot(annoying, aes(1, 2)) +
  geom_point()
```
  

  + Creating a new column called 3 which is 2 divided by 1.
  
```{r}
annoying3 <- annoying %>%
  mutate(`3` = `2`/`1`)
annoying3
```
  

  + Renaming the columns to one, two and three.
  
```{r}

# just renames the columns but keeps all
annoying3 %>%
  rename(one = `1`,
         two = `2`,
         three = `3`)

# another way but it drops other columns
annoying3 %>%
  select(one = `1`,
         two = `2`,
         three = `3`)
```
  
  
5. What does tibble::enframe() do? When might you use it?

From help using ?enframe:
  converts named atomic vectors or lists to one- or two-column data frames. For a list, the result will be a nested tibble with a column of type list. For unnamed vectors, the natural sequence is used as name column.
  
  When would we use this? Perhaps to convert lists to tibble. Convert two vectors into data frame. Maybe not application to our large datasets.
  
```{r}
enframe(mtcarstib$mpg)

```
  

6. What option controls how many additional column names are printed at the footer of a tibble?

```{r}

# normal tibble defaults
print(nycflights13::flights)

# with max columns defined 
print(nycflights13::flights, max_extra_cols = 5)

# global options
#(tibble.print_min = Inf) # option to print everything like in data frames
options(tibble.print_max = 50, tibble.print_min = 10) # does this work?

# with global tibble defaults changed
print(nycflights13::flights)

# or you can change global options for printing collumns and rows of tibbles
# (tibble.print_min = Inf)
# options(tibble.print_max = n, tibble.print_min = m)

```
  
  