---
title: "11 Data Import"
author: "Erika Anderson"
date: "May 21, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load library
library(tidyverse)

```

A group of DFO coworkers decided to work together to compete the exercises in [R for Data Science](https://r4ds.had.co.nz/) by Garrett Grolemund and Hadley Wickham. Many of the solutions to the exercises are available online, with the most [extensive version](https://jrnold.github.io/r4ds-exercise-solutions/index.html) that I have found so far by Jeffrey Arnold on Git Hub.

# 11.2.2 Exercises

1. What function would you use to read a file where fields were separated with
“|”?

read_delim

2. Apart from file, skip, and comment, what other arguments do read_csv() and read_tsv() have in common?

read_csv and read_tsv common arguements:
    
* col_names
* col_types
* locale
* na
* quoted_na
* quote
* trm_ws
* n_max
* show_progress
* skip_empty_rows


3. What are the most important arguments to read_fwf()?

file
col_positions: fwf_widths() or fwf_positions()



4. Sometimes strings in a CSV file contain commas. To prevent them from causing problems they need to be surrounded by a quoting character, like " or '. By convention, read_csv() assumes that the quoting character will be ", and if you want to change it you’ll need to use read_delim() instead. What arguments do you need to specify to read the following text into a data frame?

```{r}

library(readr)
read_delim("x,y\n1,'a,b'", delim = ",", quote = "'")

```

5. Identify what is wrong with each of the following inline CSV files. What happens when you run the code?

```{r}
read_csv("a,b\n1,2,3\n4,5,6") # two column names but three columns of data so last column dropped
read_csv("a,b,c\n1,2\n1,2,3,4") # three column names but only two values in second line of data, so NA assigned
read_csv("a,b\n\"1") # two columns and one data value so NA assigned
read_csv("a,b\n1,2\na,b") # the numbers are treated as characters since the data types are disparate
read_csv("a;b\n1;3") # semi column doesn't work as a spacer but is included as character
```


# 11.3.5 Exercises

1. What are the most important arguments to locale()?

date_names - language code as string from date_names_langs()  
date_format - has default %AD  
time_format -  has default %AT  
decimal_mark -  default is ,  
grouping_mark - default is non, but can change from , to ' or space  
tz -  time zone ; default is UTC  
encoding - default UTF-8  


2. What happens if you try and set decimal_mark and grouping_mark to the same character? What happens to the default value of grouping_mark when you set decimal_mark to “,”? What happens to the default value of decimal_mark when you set the grouping_mark to “.”?

```{r}
#locale(decimal_mark = ".", grouping_mark = ".")
# Error: `decimal_mark` and `grouping_mark` must be different

```


3. I didn’t discuss the date_format and time_format options to locale(). What do they do? Construct an example that shows when they might be useful.

They allow you to give a default format to locale(). The date_format default %AD uses an automatic data parser that recognizes Y-m-d or Y/m/d. It may be useful to change if the data is in a non-standard date format such as d-m-Y. The time_default is %AT which uses an automatic time parser the uses H:M with optional seconds and am/pm. This could be changed if the time format was in a non-stand time format such as 19h30.

```{r}

parse_date("May 21 2019")
parse_date("May 21 2019", "%B %d %Y")
parse_date("May 21 2019", locale = locale("en", date_format = "%B %d %Y"))
# Why would date_format be preferred over the shorter version with format defined?

```


4. If you live outside the US, create a new locale object that encapsulates the settings for the types of file you read most commonly.

```{r}

# see the existing locale
locale()

# create a different locale
diff_locale <- locale(date_format = "%B %d, %Y")

# see the resulting locale
diff_locale

# test the new locale with month before day
parse_date("May 21, 2019", locale = diff_locale)


```


5. What’s the difference between read_csv() and read_csv2()?

read_csv2 uses a semi colon as a deliminator so commas may be used for decimal points for European users.

6. What are the most common encodings used in Europe? What are the most common encodings used in Asia? Do some googling to find out.

Top 3 Worldwide   
1. UTF-8 93.5%  
2. ISO-8859-1 3.2%  
3. Windows-1251 0.9%   

Europe ISO-8859 many versions

Asia  Guobiao GB 2312, GBK, GB 18030


7. Generate the correct format string to parse each of the following dates and times:

```{r}
d1 <- "January 1, 2010"
parse_date(d1, "%B %d, %Y")

d2 <- "2015-Mar-07"
parse_date(d2, "%Y-%b-%d")

d3 <- "06-Jun-2017"
parse_date(d3, "%d-%b-%Y")

d4 <- c("August 19 (2015)", "July 1 (2015)")
parse_date(d4, "%B %d (%Y)")

d5 <- "12/30/14" # Dec 30, 2014
parse_date(d5, "%m/%d/%y")

t1 <- "1705"
parse_time(t1, "%H%M")

t2 <- "11:15:10.12 PM"
parse_time(t2, "%I:%M:%OS %p")
```

