---
title: "12 Tidy Data"
author: "Erika Anderson"
date: "May 21, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load library
library(tidyverse)

```

A group of DFO coworkers decided to work together to complete the exercises in [R for Data Science](https://r4ds.had.co.nz/) by Garrett Grolemund and Hadley Wickham. 

# 12.2.1 Exercises

1. Using prose, describe how the variables and observations are organised in each of the sample tables.

table1 - There are four columns (country, year, cases, population), Each observation is in a row.  
table2 - One of the columns called type indicates whether the results are cases or population. This means that there are two rows for each country row combination.  
table3 - One column called rate has the number of cases and the population within one field.  
table4 - There are two tables a and b that contain the cases and population data seperately. 


2. Compute the rate for table2, and table4a + table4b. 

You will need to perform four operations for each table:
1. Extract the number of TB cases per country per year.
2. Extract the matching population per country per year.
3. Divide cases by population, and multiply by 10000.
4. Store back in the appropriate place.

```{r}
# table 2
table2

# Extract the number of TB cases per country per year.
# Extract the matching population per country per year.
spread(table2, type, count) %>%

# Divide cases by population, and multiply by 10000.
# Store back in the appropriate place.
  mutate(., rate = cases/population * 10000)

# table 4a
table4a

# gather year columns into observations
table4a_mod <- gather(table4a, "year", "cases", 2:3)

# table 4b
table4b

# gather year columns into observations
table4b_mod <- gather(table4b, "year", "population", 2:3)

table4 <- table4a_mod %>%
  full_join(., table4b_mod, by = c("country", "year")) %>%
  mutate(., rate = cases/population * 10000)
table4

```


Which representation is easiest to work with? Which is hardest? Why?

table2 is faster. I find tables 4a and 4b more intuitive, although there is repetition to gather the two tables

4. Recreate the plot showing change in cases over time using table2 instead of table1. What do you need to do first?

```{r}
table2

# tidy data
table2_tidy <- spread(table2, type, count)
  
  # graph cases by year
  ggplot(data = table2_tidy,
         aes(year, cases, color = country)) +
  geom_point() +
  geom_line()
  
  # or a shorter version is
  # tidy data
spread(table2, type, count) %>%
  
  # graph cases by year
  ggplot(data = .,
         aes(year, cases, color = country)) +
  geom_point() +
  geom_line()


```



# 12.3.3 Exercises

1. Why are gather() and spread() not perfectly symmetrical?
Carefully consider the following example:
(Hint: look at the variable types and think about column names.)

Year is stored as a character after spreading and gathering since the column name is stored as a character.    
The order of the columns also changes from year-half-return to half-year-return.

```{r}
stocks <- tibble(
  year   = c(2015, 2015, 2016, 2016),
  half  = c(   1,    2,     1,    2),
  return = c(1.88, 0.59, 0.92, 0.17))
stocks

stocks %>%
  spread(year, return)

stocks %>%
  spread(year, return) %>%
  gather("year", "return", `2015`:`2016`)
```

2. Why does this code fail?
The names of the columns are numbers which are non-standard so must be written with back ticks.

```{r}
table4a

# table4a %>%
#   gather(1999, 2000, key = "year", value = "cases")
# #> Error in inds_combine(.vars, ind_list): Position must be between 0 and n

table4a %>%
  gather("1999", "2000", key = "year", value = "cases")

```

3. Why does spreading this tibble fail? How could you add a new column to fix the problem?

This fails because Philip Woods has two ages. There would need to be a different column to identify the date of the measurement.

```{r}
people <- tribble(
  ~name,             ~key,    ~value,
  #-----------------|--------|------
  "Phillip Woods",   "age",       45,
  "Phillip Woods",   "height",   186,
  "Phillip Woods",   "age",       50,
  "Jessica Cordero", "age",       37,
  "Jessica Cordero", "height",   156
)

people

```


4. Tidy the simple tibble below. Do you need to spread or gather it? What are the variables?

```{r}
preg <- tribble(
  ~pregnant, ~male, ~female,
  "yes",     NA,    10,
  "no",      20,    12
)
preg

# wrong way
preg %>%
spread(key = "pregnant", value = "female")

# right way
preg %>%
gather(male, female, key = "gender", value = "Number")
```

# 12.3.3 Exercises

1. What do the extra and fill arguments do in separate()? Experiment with the various options for the following two toy datasets.

extra controls what happens if there are too many pieces. The default is to warn, then drop the values. If you use extra = "merge", the remainder of the longer value is kept in the last column.  

fill controls what happens if there are not enough pieces. The default is to warn and fill from right, leaving the NA on the left edge. You can change this by using fill = "left" to get an NA on the left and fill from the left to right.  


```{r}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j"))

# original data frame with no options
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"))

# set extra option to merge
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"), extra = "merge")

# original toy data frame with one column with less
tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"))

# set fill option to left
tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"), fill = "left")
```


2. Both unite() and separate() have a remove argument. What does it do? Why would you set it to FALSE?

The remove = TRUE options means the original column will be removed from the resulting data frame. You can set it to FALSE to remove the original field.

3. Compare and contrast separate() and extract(). Why are there three variations of separation (by position, by separator, and with groups), but only one unite?

separate() and extract() both split a column into multiple columns using a data frame and column as input. It looks like separate require a common item to split the column, whereas extract can uses regex to determine where to break the columns up so it is more flexible.

unite() puts columns together so there are less options than with separate().

# 12.5.1 Exercises

1. Compare and contrast the fill arguments to spread() and complete().

The fill option in complete() takes the last value in that column and using it to fill in the missing values. The fill option in spread() tells the function what to replace missing values with.

Spread only provides missing values within dataset but may not infill implicit missing values that are not present.

Complete has a list option for fill but spread only takes a value as input.

```{r }
# try with years and pattern to fill

```

2. What does the direction argument to fill() do?

The direction arguement tells the direction in which to fill missing values, either "down" (default) or "up".

# 12.6.1 Exercises

1. In this case study I set na.rm = TRUE just to make it easier to check that we had the correct values. Is this reasonable? Think about how missing values are represented in this dataset. Are there implicit missing values? What’s the difference between an NA and zero?

The difference between an NA and zero is that NA means nothing was counted, whereas zero means there were no cases. Given the data set, it may be reasonable to assume that NAs may be dropped as not counted because zeros are explicitly recorded.

```{r}
# original dataset
who

# gather value columns into key value pairs
who1 <- who %>% 
  gather(new_sp_m014:newrel_f65, key = "key", value = "cases", na.rm = TRUE)
who1

# add underscore for newrel
who2 <- who1 %>% 
  mutate(key = stringr::str_replace(key, "newrel", "new_rel"))
who2

# split the key column into sections
who3 <- who2 %>% 
  separate(key, c("new", "type", "sexage"), sep = "_")
who3

# remove redundant country columns
who4 <- who3 %>% 
  select(-new, -iso2, -iso3)

# separate sex and age into two columns
who5 <- who4 %>%
  separate(sexage, c("sex", "age"), sep = 1)
who5

```


2. What happens if you neglect the mutate() step? (mutate(key = stringr::str_replace(key, "newrel", "new_rel")))

A warning is generated and NAs are inserted

```{r}

# split the key column into sections
who_test_ <- who1 %>% 
  separate(key, c("new", "type", "sexage"), sep = "_")
who_test

```


3. I claimed that iso2 and iso3 were redundant with country. Confirm this claim.

No duplicates if distinct cases, insluding iso2 and iso3 are included.

```{r}
who_test2 <- who3 %>%
  select(country, iso2, iso3) %>%
  distinct()
who_test2

# check for duplicates
anyDuplicated(who_country)

who_iso2 <- who_test2 %>%
  select(iso2)
  
# check for duplicates
anyDuplicated(who_iso2)

who_iso3 <- who_test2 %>%
  select(iso3)
  
# check for duplicates
anyDuplicated(who_iso3)

# check for total combinations and compare to nrow(who_country)
n_distinct(who_country)

```


4. For each country, year, and sex compute the total number of cases of TB. Make an informative visualisation of the data.

```{r}
# tidy data
who5

# summarize the counts
who_sum <- who5 %>%
  group_by(country, year, sex) %>%
  summarise(totalTB = sum(cases)) %>%
  ungroup()
who_sum

# select top countries to display
who_top9 <- who_sum %>%
  group_by(country) %>%
  summarize(totalCountry = sum(totalTB)) %>%
  arrange(., desc(totalCountry)) %>%
  slice(1:9) %>%
  ungroup() %>%
  pull(country)
who_top9

# take top from summed data
who_top <- who_sum %>%
  filter(country %in% who_top9)
  
# graph it
ggplot(who_top, aes(year, totalTB, color = sex)) +
  facet_wrap(~country) +
  geom_point() +
  geom_line() +
  labs(title = "Countries with Highest Recorded TB Counts") +
  theme_bw()

```

