---
title: "12 Tidy Data"
author: "Erika Anderson"
date: "May 21, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load library
library(tidyverse)

```

A group of DFO coworkers decided to work together to compete the exercises in [R for Data Science](https://r4ds.had.co.nz/) by Garrett Grolemund and Hadley Wickham. 

# 12.2.1 Exercises

1. Using prose, describe how the variables and observations are organised in each of the sample tables.

table1 - There are four columns (country, year, cases, population), Each observation is in a row.  
table2 - One of the columns called type indicates whether the results are cases or population. This means that there are two rows for each country row combination.  
table3 - One column called rate has the number of cases and the population within one field.  
table4 - There are two tables a and b that contain the cases and population data seperately. 


2. Compute the rate for table2, and table4a + table4b. 

You will need to perform four operations for each table:
1. Extract the number of TB cases per country per year.
2. Extract the matching population per country per year.
3. Divide cases by population, and multiply by 10000.
4. Store back in the appropriate place.

```{r}
# table 2
table2

# Extract the number of TB cases per country per year.
# Extract the matching population per country per year.
spread(table2, type, count) %>%

# Divide cases by population, and multiply by 10000.
# Store back in the appropriate place.
  mutate(., rate = cases/population * 10000)

# table 4a
table4a

# gather year columns into observations
table4a_mod <- gather(table4a, "year", "cases", 2:3)

# table 4b
table4b

# gather year columns into observations
table4b_mod <- gather(table4b, "year", "population", 2:3)

table4 <- table4a_mod %>%
  full_join(., table4b_mod, by = c("country", "year")) %>%
  mutate(., rate = cases/population * 10000)
table4

```


Which representation is easiest to work with? Which is hardest? Why?

table2 is faster. I find tables 4a and 4b more intuitive, although there is repetition to gather the two tables

4. Recreate the plot showing change in cases over time using table2 instead of table1. What do you need to do first?

```{r}
table2

# tidy data
table2_tidy <- spread(table2, type, count)
  
  # graph cases by year
  ggplot(data = table2_tidy,
         aes(year, cases, color = country)) +
  geom_point() +
  geom_line()
  
  # or a shorter version is
  # tidy data
spread(table2, type, count) %>%
  
  # graph cases by year
  ggplot(data = .,
         aes(year, cases, color = country)) +
  geom_point() +
  geom_line()


```



# 12.3.3 Exercises

1. Why are gather() and spread() not perfectly symmetrical?
Carefully consider the following example:
(Hint: look at the variable types and think about column names.)

```{r}
# stocks <- tibble(
#   year   = c(2015, 2015, 2016, 2016),
#   half  = c(   1,    2,     1,    2),
#   return = c(1.88, 0.59, 0.92, 0.17)
# )
# stocks %>% 
#   spread(year, return) %>% 
#   gather("year", "return", `2015`:`2016`)
```

2. Why does this code fail?

```{r}
# table4a %>% 
#   gather(1999, 2000, key = "year", value = "cases")
# #> Error in inds_combine(.vars, ind_list): Position must be between 0 and n
```

3. Why does spreading this tibble fail? How could you add a new column to fix the problem?

```{r}
people <- tribble(
  ~name,             ~key,    ~value,
  #-----------------|--------|------
  "Phillip Woods",   "age",       45,
  "Phillip Woods",   "height",   186,
  "Phillip Woods",   "age",       50,
  "Jessica Cordero", "age",       37,
  "Jessica Cordero", "height",   156
)
```


4. Tidy the simple tibble below. Do you need to spread or gather it? What are the variables?

```{r}
preg <- tribble(
  ~pregnant, ~male, ~female,
  "yes",     NA,    10,
  "no",      20,    12
)
```

# 12.3.3 Exercises

1. What do the extra and fill arguments do in separate()? Experiment with the various options for the following two toy datasets.

```{r}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"))

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"))
```


2. Both unite() and separate() have a remove argument. What does it do? Why would you set it to FALSE?

3. Compare and contrast separate() and extract(). Why are there three variations of separation (by position, by separator, and with groups), but only one unite?

# 12.5.1 Exercises

1. Compare and contrast the fill arguments to spread() and complete().

2. What does the direction argument to fill() do?

# 12.6.1 Exercises

1. In this case study I set na.rm = TRUE just to make it easier to check that we had the correct values. Is this reasonable? Think about how missing values are represented in this dataset. Are there implicit missing values? What’s the difference between an NA and zero?

2. What happens if you neglect the mutate() step? (mutate(key = stringr::str_replace(key, "newrel", "new_rel")))

3. I claimed that iso2 and iso3 were redundant with country. Confirm this claim.

4. For each country, year, and sex compute the total number of cases of TB. Make an informative visualisation of the data.