---
title: "14 Strings"
author: "Erika Anderson"
date: "July 5, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load library
library(tidyverse)
library(stringr)

```

A group of DFO coworkers decided to work together to compete the exercises in [R for Data Science](https://r4ds.had.co.nz/) by Garrett Grolemund and Hadley Wickham. 

# 14.2.5 Exercises

1. In code that doesn’t use stringr, you’ll often see paste() and paste0(). What’s the difference between the two functions? What stringr function are they equivalent to? How do the functions differ in their handling of NA?

paste and paste0 differ in that paste uses a separator and paste0 leaves no blank or separator.  

They are similar to the str_c() function, with sep for paste and without sep for paste0. Str_c is better since it maintains NAs instead of making them character NAs.


2. In your own words, describe the difference between the sep and collapse arguments to str_c().

collapse creates one string from the input, even it has a separator.  
sep create multiple separate values in multiple strings.

```{r}

str_c(c("x", "y", "z"), collapse = ", ")

str_c(c("x", "y", "z"), sep = ", ")

```


3. Use str_length() and str_sub() to extract the middle character from a string. What will you do if the string has an even number of characters?

str_sub rounds up if there is an even number of characters.

```{r}

x <- c("Apple", "Banana", "Pear")

str_sub(x, (str_length(x)/2), (str_length(x)/2))

```


4. What does str_wrap() do? When might you want to use it?

str_wrap wraps strings into formatted paragraphs.  
Not sure when I would use it since I prefer Rmarkdown.


5. What does str_trim() do? What’s the opposite of str_trim()?

str_trim removes whitespace from strings which helps if you are using str_length to pull a specific character from the string. The opposite is str_pad() to add whitespace or any character.

6. Write a function that turns (e.g.) a vector c("a", "b", "c") into the string a, b, and c. Think carefully about what it should do if given a vector of length 0, 1, or 2.

```{r}

vec_to_string <- function(vec, sep = ",") {
  n <- length(vec)
  
  if (n == 0) {""}
  else if (n == 1) {vec}
  else if (n == 2) {str_c(vec[[1]], "and", vec[[2]], sep = " ")}
  else {
     # commas after all n - 1 elements
    not_last <- str_c(vec[seq_len(n - 1)], sep)
    # prepend "and" to the last element
    last <- str_c("and", vec[[n]], sep = " ")
    # combine parts with spaces
    str_c(c(not_last, last), collapse = " ")
  }
}

testzero <- ""
testone <- "a"
testtwo <- c("a","b")
testthree <- c("a", "b", "c")

vec_to_string(testzero)
vec_to_string(testone)
vec_to_string(testtwo)
vec_to_string(testthree)
```

# 14.3 Matching patterns with regular expressions

## 14.3.1.1 Exercises

1. Explain why each of these strings don’t match a \: "\", "\\", "\\\".

One "\" is an escape from R code
Two "\\" is an escape from R code and regex
Three "\\\" is an escape from R code, regex, and a string making the regex


2. How would you match the sequence "'\?

```{r}

x <- "\"'\\"
str_view(x, "\"'\\\\")

```


3. What patterns will the regular expression \..\..\.. match? How would you represent it as a string?

A dot represents any character except a new line, but a back slash dot represents an actual dot. So this example will find any three characters seperated by actual dots.

```{r}

x <- c(".a.b.c", ".a.b", ".....")
str_view_all(x, c("\\..\\..\\.."))

```

14.3.2.1 Exercises

1. How would you match the literal string "$^$"?

```{r}

str_view("$^$", "\\$\\^\\$$")

```

2. Given the corpus of common words in stringr::words, create regular expressions that find all words that:

Start with “y”.

```{r}
str_view_all(stringr::words, "^y", match = TRUE)
```

End with “x”

```{r}
str_view_all(stringr::words, "x$", match = TRUE)
```

Are exactly three letters long. (Don’t cheat by using str_length()!)

```{r}
str_view_all(stringr::words, "^...$", match = TRUE)
```

Have seven letters or more.

```{r}
str_view_all(stringr::words, "^.......$", match = TRUE)
```


## 14.3.3.1 Exercises
1. Create regular expressions to find all words that:

Start with a vowel.
```{r}
str_view_all(stringr::words, "^[aeiou]", match = TRUE)
```

That only contain consonants. (Hint: thinking about matching “not”-vowels.)
```{r}
str_view_all(stringr::words, "^[^aeiou]+$", match = TRUE)
```

End with ed, but not with eed.
```{r}
str_view_all(stringr::words, "[^e]ed$", match = TRUE)
```

End with ing or ise.
```{r}
str_view_all(stringr::words, "ing$|ise$", match = TRUE)
```

2. Empirically verify the rule “i before e except after c”.

No - there are always exceptions in English.

```{r}

str_view_all(stringr::words, "cie", match = TRUE)

```


3. Is “q” always followed by a “u”? 

Yes in stringr words dataset.
```{r}

str_view_all(stringr::words, "q[^u]", match = TRUE)

```


4. Write a regular expression that matches a word if it’s probably written in British English, not American English.

I used "our: as an example as American English uses "or".

```{r}

str_view_all(stringr::words, "..+our$", match = TRUE)

```



5. Create a regular expression that will match telephone numbers as commonly written in your country.
```{r}

str_view_all(c("250-123-7456", "25-45-62-58"), "\\d\\d\\d-\\d\\d\\d-\\d\\d\\d\\d", match = TRUE)

```

## 14.3.4.1 Exercises
1. Describe the equivalents of ?, +, * in {m,n} form.

? = {0,1}  

+ = {1,}  

* = {0,}  

2. Describe in words what these regular expressions match: (read carefully to see if I’m using a regular expression or a string that defines a regular expression.)

^.*$ = anchored to start, zero to more characters, before end  

"\\{.+\\}" = any string with one or more characters surrounded by curly brackets   

\d{4}-\d{2}-\d{2} = four digits, dash, two digits, dash, two digits  

"\\\\{4}" = back slash at least 4 times  


3. Create regular expressions to find all words that:

Start with three consonants.
```{r}
str_view_all(stringr::words, "^[^aeiou]{3}", match = TRUE)
```


Have three or more vowels in a row.
```{r}
str_view_all(stringr::words, "[aeiou]{3,}", match = TRUE)
```

Have two or more vowel-consonant pairs in a row.
```{r}
str_view_all(stringr::words, "([aeiou][^aeiou]){2,}", match = TRUE)

```


4. Solve the beginner regexp crosswords at https://regexcrossword.com/challenges/beginner.

Started with tuorial so I could figure it out.

## 14.3.5 Grouping and Backreferences Exercises

1. Describe, in words, what these expressions will match:

(.)\1\1 Matches one character that repeats itself three times

"(.)(.)\\2\\1" Not sure how reverse grouped back references work, but Jrnold says "A pair of characters followed by the same pair of characters in reversed order. E.g. "abba"."
See https://jrnold.github.io/r4ds-exercise-solutions/strings.html#matching-patterns-with-regular-expressions


(..)\1 any two characters repeated once

"(.).\\1.\\1" Macthes a character followed by any character, followed by itself again, then any character, then itself. For example: 19181

"(.)(.)(.).*\\3\\2\\1" Again, not sure why there are 3,2,1 groups and how it gets reversed.

Jrnold says "Three characters followed by zero or more characters of any kind followed by the same three characters but in reverse order. E.g. "abcsgasgddsadgsdgcba" or "abccba" or "abc1cba"."

2. Construct regular expressions to match words that:

Start and end with the same character.
```{r}
str_view_all(stringr::words, "^(.)(.*\\1$)", match = TRUE)
```


Contain a repeated pair of letters (e.g. “church” contains “ch” repeated twice.)
```{r}
str_view_all(stringr::words, "([a-z][a-z]).*\\1", match = TRUE)
```


Contain one letter repeated in at least three places (e.g. “eleven” contains three “e”s.)
```{r}
str_view_all(stringr::words, "([a-z]).*\\1.*\\1", match = TRUE)
```

