---
title: "14 Strings"
author: "Erika Anderson"
date: "July 5, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load library
library(tidyverse)
library(stringr)

```

A group of DFO coworkers decided to work together to compete the exercises in [R for Data Science](https://r4ds.had.co.nz/) by Garrett Grolemund and Hadley Wickham. 

# 14.2.5 Exercises

1. In code that doesn’t use stringr, you’ll often see paste() and paste0(). What’s the difference between the two functions? What stringr function are they equivalent to? How do the functions differ in their handling of NA?

paste and paste0 differ in that paste uses a separator and paste0 leaves no blank or separator.  

They are similar to the str_c() function, with sep for paste and without sep for paste0. Str_c is better since it maintains NAs instead of making them character NAs.


2. In your own words, describe the difference between the sep and collapse arguments to str_c().

collapse creates one string from the input, even it has a separator.  
sep create multiple separate values in multiple strings.

```{r}

str_c(c("x", "y", "z"), collapse = ", ")

str_c(c("x", "y", "z"), sep = ", ")

```


3. Use str_length() and str_sub() to extract the middle character from a string. What will you do if the string has an even number of characters?

str_sub rounds up if there is an even number of characters.

```{r}

x <- c("Apple", "Banana", "Pear")

str_sub(x, (str_length(x)/2), (str_length(x)/2))

```


4. What does str_wrap() do? When might you want to use it?

str_wrap wraps strings into formatted paragraphs.  
Not sure when I would use it since I prefer Rmarkdown.


5. What does str_trim() do? What’s the opposite of str_trim()?

str_trim removes whitespace from strings which helps if you are using str_length to pull a specific character from the string. The opposite is str_pad() to add whitespace or any character.

6. Write a function that turns (e.g.) a vector c("a", "b", "c") into the string a, b, and c. Think carefully about what it should do if given a vector of length 0, 1, or 2.

```{r}

vec_to_string <- function(vec, sep = ",") {
  n <- length(vec)
  
  if (n == 0) {""}
  else if (n == 1) {vec}
  else if (n == 2) {str_c(vec[[1]], "and", vec[[2]], sep = " ")}
  else {
     # commas after all n - 1 elements
    not_last <- str_c(vec[seq_len(n - 1)], sep)
    # prepend "and" to the last element
    last <- str_c("and", vec[[n]], sep = " ")
    # combine parts with spaces
    str_c(c(not_last, last), collapse = " ")
  }
}

testzero <- ""
testone <- "a"
testtwo <- c("a","b")
testthree <- c("a", "b", "c")

vec_to_string(testzero)
vec_to_string(testone)
vec_to_string(testtwo)
vec_to_string(testthree)
```

# 14.3 Matching patterns with regular expressions

## 14.3.1.1 Exercises

1. Explain why each of these strings don’t match a \: "\", "\\", "\\\".

One "\" is an escape from R code
Two "\\" is an escape from R code and regex
Three "\\\" is an escape from R code, regex, and a string making the regex


2. How would you match the sequence "'\?

```{r}

x <- "\"'\\"
writeLines(x)

# another way to make it
y <- paste0("\"", "'", "\\")
writeLines(y)

str_view(x, "\"'\\\\")

```


3. What patterns will the regular expression \..\..\.. match? How would you represent it as a string?

A dot represents any character except a new line, but a back slash dot represents an actual dot. So this example will find any three characters separated by actual dots.

```{r}

x <- c(".a.b.c", ".a.b", ".....")
str_view_all(x, c("\\..\\..\\.."))

```

14.3.2.1 Exercises

1. How would you match the literal string "$^$"?

```{r}

str_view("$^$", "\\$\\^\\$$")

```

2. Given the corpus of common words in stringr::words, create regular expressions that find all words that:

Start with “y”.

```{r}
str_view_all(stringr::words, "^y", match = TRUE)
```

End with “x”

```{r}
str_view_all(stringr::words, "x$", match = TRUE)
```

Are exactly three letters long. (Don’t cheat by using str_length()!)

```{r}
str_view_all(stringr::words, "^[a-z]{3}$", match = TRUE)

```

Have seven letters or more.

```{r}
str_view_all(stringr::words, "^.......$", match = TRUE)
str_view_all(stringr::words, "^[:alpha:]{7}$", match = TRUE) # matches uppper and lower case
```


## 14.3.3.1 Exercises
1. Create regular expressions to find all words that:

Start with a vowel.
```{r}
str_view_all(stringr::words, "^[aeiou]", match = TRUE) # only matches lower cases
str_view_all(stringr::words, "^[aeiouAEIOU]", match = TRUE) # matches both cases
```

That only contain consonants. (Hint: thinking about matching “not”-vowels.)
```{r}
str_view_all(stringr::words, "^[^aeiou]+$", match = TRUE)
str_view_all(stringr::words, "^[^aeiouAEIOU]+$", match = TRUE) # matches both cases
```

End with ed, but not with eed.
```{r}
str_view_all(stringr::words, "[^e]ed$", match = TRUE)
```

End with ing or ise.
```{r}
str_view_all(stringr::words, "ing$|ise$", match = TRUE)
```

2. Empirically verify the rule “i before e except after c”.

No - there are always exceptions in English.

```{r}

str_view_all(stringr::words, "cie", match = TRUE)
str_view_all(stringr::words, "[^c]ei", match = TRUE)

```


3. Is “q” always followed by a “u”? 

Yes in stringr words dataset.
```{r}

str_view_all(stringr::words, "q[^u]", match = TRUE)

```


4. Write a regular expression that matches a word if it’s probably written in British English, not American English.

I used "our: as an example as American English uses "or".

```{r}

str_view_all(stringr::words, "..+our$", match = TRUE)
str_view_all(stringr::words, ".{2,}our$", match = TRUE)

```



5. Create a regular expression that will match telephone numbers as commonly written in your country.
```{r}

str_view_all(c("250-123-7456", "25-45-62-58"), "\\d\\d\\d-\\d\\d\\d-\\d\\d\\d\\d", match = TRUE)
str_view_all(c("250-123-7456", "25-45-62-58"), "\\d{3}-\\d{3}-\\d{4}", match = TRUE)

```

## 14.3.4.1 Exercises
1. Describe the equivalents of ?, +, * in {m,n} form.

? = {0,1}  

+ = {1,}  

* = {0,}  

2. Describe in words what these regular expressions match: (read carefully to see if I’m using a regular expression or a string that defines a regular expression.)

^.*$ = anchored to start, zero to more characters, before end  

"\\{.+\\}" = any string with one or more characters surrounded by curly brackets   

\d{4}-\d{2}-\d{2} = four digits, dash, two digits, dash, two digits  

"\\\\{4}" = back slash at least 4 times  


3. Create regular expressions to find all words that:

Start with three consonants.
```{r}
str_view_all(stringr::words, "^[^aeiouy]{3}", match = TRUE) #MO wants y included
```


Have three or more vowels in a row.
```{r}
str_view_all(stringr::words, "[aeiou]{3,}", match = TRUE) # include upper case, if needed
str_view_all(stringr::words, "[aeiouy]{3,}", match = TRUE) # MO wants y included
```

Have two or more vowel-consonant pairs in a row.
```{r}
str_view_all(stringr::words, "([aeiou][^aeiou]){2,}", match = TRUE)

```


4. Solve the beginner regexp crosswords at https://regexcrossword.com/challenges/beginner.

Started with tutorial, so I could figure it out.

## 14.3.5 Grouping and Backreferences Exercises

1. Describe, in words, what these expressions will match:

(.)\1\1 Matches one character that repeats itself three times

"(.)(.)\\2\\1"  Jrnold says "A pair of characters followed by the same pair of characters in reversed order. E.g. "abba"."
See https://jrnold.github.io/r4ds-exercise-solutions/strings.html#matching-patterns-with-regular-expressions


(..)\1 any two characters repeated once ie. bebe

"(.).\\1.\\1" Matches a character followed by any character, followed by itself again, then any character, then itself. For example: 1a1b1

"(.)(.)(.).*\\3\\2\\1" ie. abclalalacba

Jrnold says "Three characters followed by zero or more characters of any kind followed by the same three characters but in reverse order. E.g. "abcsgasgddsadgsdgcba" or "abccba" or "abc1cba"."

2. Construct regular expressions to match words that:

Start and end with the same character.
```{r}
str_view_all(stringr::words, "^(.).*\\1$", match = TRUE)
```


Contain a repeated pair of letters (e.g. “church” contains “ch” repeated twice.)
```{r}
str_view_all(stringr::words, "([a-z][a-z]).*\\1", match = TRUE) # lower case only in this example
```


Contain one letter repeated in at least three places (e.g. “eleven” contains three “e”s.)
```{r}
str_view_all(stringr::words, "([a-z]).*\\1.*\\1", match = TRUE)
```

# 14.4.1 Detect matches

1. For each of the following challenges, try solving it by using both a single regular expression, and a combination of multiple str_detect() calls.

- Find all words that start or end with x.

```{r}

str_subset(words, "^x|x$")

start_x <- str_detect(words, "^x")
end_x <- str_detect(words, "x$")
words[start_x | end_x]

```


- Find all words that start with a vowel and end with a consonant.

```{r}
start_vowel <- str_detect(words, "^[aeiou]")
end_consonant <- str_detect(words, "[^aeiou]$")
words[start_vowel & end_consonant]

str_subset(words, "^[aeiou].+[^aeiou]$")
```


- Are there any words that contain at least one of each different vowel?

```{r}

# not sure how to do it with with one expression since the order changes so it would be many combinations

# the simpler way is
words[str_detect(words, "a") &
      str_detect(words, "e") &
      str_detect(words, "i") &
      str_detect(words, "o") &
      str_detect(words, "u")]
```


2. What word has the highest number of vowels? What word has the highest proportion of vowels? (Hint: what is the denominator?)

```{r}

# use highest value of vowels to extract from the data
words[max(str_count(words, "[aeiou]"))]

# highest proportion of consonants
words[max(str_count(words, "[^aeiou]") / str_length(words))]
```



# 14.4.2 Extract matches

1. In the previous example, you might have noticed that the regular expression matched “flickered”, which is not a colour. Modify the regex to fix the problem.

First tried the pattern with one space. This omitted the colors at the end of a sentence. Looked up the word boundary symbol backslash b, and needed to escape.

```{r}

# make color pattern
colours <- str_c("\\b", c("red", "orange", " +yellow +", " +green +", " +blue", " +purple +"), "\\b")
colour_match <- str_c(colours, collapse = "|")
colour_match

# find sentence with issue
more <- sentences[str_count(sentences, colour_match) > 1]
str_view_all(more, colour_match)
```


2. From the Harvard sentences data, extract:

- The first word from each sentence.
```{r}
head(str_extract(sentences, "[a-zA-Z]+"))
```

- All words ending in ing.
```{r}

# make pattern that matches words ending in ing
pattern <- "\\b[a-zA-Z]+ing\\b"

# detect which sentences have ing words
sentences_ing <- str_detect(sentences, pattern)

# extract unique words from those sentences
unique(unlist(str_extract_all(sentences[sentences_ing], pattern)))
```

- All plurals.

This is an oversimplification since not all plurals end in s and some words ending in s are not plural. Not sure how to do it without a list of plural words given.

```{r}
# make a pattern to represent words ending in s
pattern <- "\\b[a-zA-Z]+s\\b"

# detect which sentences that have words ending in s
sentences_pl <- str_detect(sentences, pattern)

# extract plural words from those sentences
unique(unlist(str_extract_all(sentences[sentences_pl], pattern)))
```


# 14.4.3 Grouped Matches

1. Find all words that come after a “number” like “one”, “two”, “three” etc. Pull out both the number and the word.

```{r}
pattern <- "(one|two|three|four|five|six|seven|eight|nine|ten) (\\w*\\b)"
subsetSentences <- (str_subset(sentences, pattern))

str_match(subsetSentences, pattern)

# other way
sentences_tibb <- tibble(sentence_col = subsetSentences)

sentences_tibb %>%
  tidyr::extract(sentence_col, c("number", "item"), pattern, remove = FALSE) 

```


2. Find all contractions. Separate out the pieces before and after the apostrophe.

```{r}

pattern <- "(\\b[a-zA-Z]+)(')([a-zA-z]\\b)"
subsetSentences <- (str_subset(sentences, pattern))

str_match(subsetSentences, pattern)

# other way
sentences_tibb <- tibble(sentence_col = subsetSentences)

sentences_tibb %>%
  tidyr::extract(sentence_col, c("firstWord", "apostrophe", "secondWord"), pattern, remove = FALSE) 

```


# 14.4.4 Replacing Matches

1. Replace all forward slashes in a string with backslashes.
```{r}
# make sentence with slashes
data <- "This is / an example of / a string / with slashes."

# replace all
# need to use four slashes for literal slash
str_replace_all(data, "/", "\\\\")

# print it as you would see it
writeLines(str_replace_all(data, "/", "\\\\"))

```


2. Implement a simple version of str_to_lower() using replace_all().


```{r}

head(sentences)

# this is the easier way
head(str_to_lower(sentences))

# need to make replacements
# I copied from website
replacements <- c(
  "A" = "a", "B" = "b", "C" = "c", "D" = "d", "E" = "e",
  "F" = "f", "G" = "g", "H" = "h", "I" = "i", "J" = "j",
  "K" = "k", "L" = "l", "M" = "m", "N" = "n", "O" = "o",
  "P" = "p", "Q" = "q", "R" = "r", "S" = "s", "T" = "t",
  "U" = "u", "V" = "v", "W" = "w", "X" = "x", "Y" = "y",
  "Z" = "z"
)

# with str_replace_all
head(str_replace_all(sentences, replacements))

```


3. Switch the first and last letters in words. Which of those strings are still words?

```{r}

# switch first and last letters
switch <- words %>%
  str_replace_all("^([a-zA-Z])(.*)([a-zA-Z])$", "\\3\\2\\1")

# which are words still
intersect(words, switch)

```


# 14.4.5 Splitting

1. Split up a string like "apples, pears, and bananas" into individual components.
```{r}
x <- c("apple", "pear", "banana")

# split based on boundary and simplfy to matrix
str_split(x, boundary(), simplify = TRUE)
```


2. Why is it better to split up by boundary("word") than " "?

The programmers have thought of all the cases that don't work with space and wrote that into boundary("word). For example, period is included with the last word in the first example, but not in the last example using boundary("word").

```{r}
head(sentences) %>%
str_split(., " ", simplify = TRUE)

head(sentences) %>%
str_split(., boundary("word"), simplify = TRUE)
```


3. What does splitting with an empty string ("") do? Experiment, and then read the documentation.

It looks like it separates everything out into single characters.The help file says: "An empty pattern, "", is equivalent to boundary("character")."

```{r}
head(sentences) %>%
str_split(., "", simplify = TRUE)


```


# 14.5.1 Other Types of Regular Expressions

1. How would you find all strings containing \ with regex() vs. with fixed()?
  
Regex is the default.Use fixed around pattern to use it. Fixed is faster but has issues with more complicated searches. The back slash does not need as many escapes since there is no regex to escape from...

```{r }
x <- c("ap\\ple", "pear", "ba\\nana")
writeLines(x)

str_subset(x, "\\\\")
str_subset(x, fixed("\\"))

```


2. What are the five most common words in sentences?

```{r}

# my way
 allWords <- str_extract_all(sentences, boundary("word"))

allWords_tibb <- as.tibble(unlist(allWords))

allWords_tibb %>%
  mutate(lowerCase = str_to_lower(value)) %>%
  group_by(lowerCase) %>%
  count() %>%
  ungroup() %>%
  arrange(-n) %>%
  slice(1:5)

# from the answer key in concise code
tibble(word = unlist(str_extract_all(sentences, boundary("word")))) %>%
  mutate(word = str_to_lower(word)) %>%
   count(word, sort = TRUE) %>%
  head(5)


```

# 14.7 stringi

1. Find the stringi functions that:

- Count the number of words.
```{r}
require(stringi)

x <- "This is an example sentence"

stri_count_words(x)
```

- Find duplicated strings.
```{r}
x <- c("this is an duplicate sentence", "This is not", "this is an duplicate sentence")

x[stri_duplicated(x)]
```

- Generate random text.
```{r}

stri_rand_strings(5, 10)

```

2. How do you control the language that stri_sort() uses for sorting?

Use locale option within stri_sort. Use locale option within stri_sort. Get list of locales possible using stri_locale_list().



