---
title: "15 Factors"
author: "Erika Anderson"
date: "August 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load library
library(tidyverse)
library(forcats)

```

A group of DFO coworkers decided to work together to complete the exercises in [R for Data Science](https://r4ds.had.co.nz/) by Garrett Grolemund and Hadley Wickham. 

# 15 Factors

# 15.3.1 Exercise
1. Explore the distribution of rincome (reported income). What makes the default bar chart hard to understand? How could you improve the plot?

```{r}
unique(gss_cat$rincome)

ggplot(gss_cat, aes(rincome)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# overlapping category labels make this hard to understand
# should order the levels by value
# combine the factors that are no answers

# try it again after completing chapter
gss_cat_mod <- gss_cat %>%
  mutate(income_simple = fct_collapse(rincome,
    Unknown = c("Not applicable", "No answer", "Don't know", "Refused"),
    "More than 25 000" = c("$25000 or more"),
    "$2000 to $25000" = c("$20000 - 24999", "$15000 - 19999", "$10000 - 14999", "$8000 to 9999",
                          "$7000 to 7999", "$6000 to 6999", "$5000 to 5999", "$4000 to 4999",
                          "$3000 to 3999" ),
    "Less than $3000" = c("$1000 to 2999", "Lt $1000")
  ))

# graph the collapsed factors
ggplot(gss_cat_mod, aes(income_simple)) +
  geom_bar() +
  labs(x = "Grouped income") 

# another way, less precise but faster
ggplot(gss_cat, aes(fct_lump(rincome, n = 5))) +
  geom_bar() +
  labs(x = "Grouped income") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))

```


2. What is the most common relig in this survey? What’s the most common partyid?

```{r}

gss_cat %>%
  count(relig) %>%
  arrange(-n)
```
```{r}

gss_cat %>%
  count(partyid) %>%
  arrange(-n)
```


3. Which relig does denom (denomination) apply to? How can you find out with a table? How can you find out with a visualisation?

Protestant

```{r}

gss_cat %>%
  group_by(relig, denom) %>%
  count() %>%
  arrange(-n)

ggplot(gss_cat, aes(denom, relig)) +
         geom_point() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))

#JN
ggplot(gss_cat, aes(relig)) +
  geom_bar(aes(fill = denom)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
  
```


# 15.4.1 Exercises
1. There are some suspiciously high numbers in tvhours. Is the mean a good summary?

No

```{r}

ggplot(gss_cat, aes(fct_reorder(relig, tvhours, mean), tvhours)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```


2. For each factor in gss_cat identify whether the order of the levels is arbitrary or principled.


```{r}

# find factors
str(gss_cat)

levels(gss_cat$marital) # arbitrary or JN/MO says principled
levels(gss_cat$race) # arbitrary or JN/MO principled
levels(gss_cat$rincome) # principled
levels(gss_cat$partyid) # arbitrary
levels(gss_cat$relig) # arbitrary and everyone agrees!
levels(gss_cat$denom) # arbitrary

```


3. Why did moving “Not applicable” to the front of the levels move it to the bottom of the plot?

The first level goes on the y axis in the first stop.
It gives an integer value of one. (JN)

```{r}

# given dataset
rincome_summary <- gss_cat %>%
  group_by(rincome) %>%
  summarise(
    age = mean(age, na.rm = TRUE),
    tvhours = mean(tvhours, na.rm = TRUE),
    n = n()
  )

# current levels of rincome
levels(rincome_summary$rincome)

# relevel
relevelled <- rincome_summary %>%
  mutate(rincome = fct_relevel(rincome, "Not applicable"))

levels(relevelled$rincome)

```


# 15.5.1 Exercises
1. How have the proportions of people identifying as Democrat, Republican, and Independent changed over time?

```{r}

gss_cat_collapse <- gss_cat %>%
  mutate(partyid2 = fct_collapse(partyid,
    other = c("No answer", "Don't know", "Other party"),
    rep = c("Strong republican", "Not str republican"),
    ind = c("Ind,near rep", "Independent", "Ind,near dem"),
    dem = c("Not str democrat", "Strong democrat")
  )) %>%
  count(year, partyid2) %>%
  group_by(year) %>%
  mutate(prop = n / sum(n))


# graph over time
ggplot(gss_cat_collapse, aes(year, n, color = fct_reorder2(partyid2, year, prop))) +
  geom_point() +
  geom_line() +
  labs(colour = "Party")

# fct_reorder2 uses y values to reorder

```


2. How could you collapse rincome into a small set of categories?

```{r}

gss_cat %>%
  mutate(rincome = fct_lump(rincome, n = 5)) %>%
  count(rincome, sort = TRUE)

```

