---
title: "16 Dates and times"
author: "Erika Anderson"
date: "September 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load libraries
library(tidyverse)
library(lubridate)
library(nycflights13)

```

A group of DFO coworkers decided to work together to complete the exercises in [R for Data Science](https://r4ds.had.co.nz/) by Garrett Grolemund and Hadley Wickham. 

# 16 Dates and times

## 16.2.4 Creating Date/times Exercises

1. What happens if you parse a string that contains invalid dates?

```{r}
ymd(c("2010-10-10", "bananas"))

```


2. What does the tzone argument to today() do? Why is it important?


3. Use the appropriate lubridate function to parse each of the following dates:

```{r}
# given dates
d1 <- "January 1, 2010"
d2 <- "2015-Mar-07"
d3 <- "06-Jun-2017"
d4 <- c("August 19 (2015)", "July 1 (2015)")
d5 <- "12/30/14" # Dec 30, 2014

```

## 16.3.4 Date-tiime components Exercises

1. How does the distribution of flight times within a day change over the course of the year?

2. Compare dep_time, sched_dep_time and dep_delay. Are they consistent? Explain your findings.

3. Compare air_time with the duration between the departure and arrival. Explain your findings. (Hint: consider the location of the airport.)

4. How does the average delay time change over the course of a day? Should you use dep_time or sched_dep_time? Why?

5. On what day of the week should you leave if you want to minimise the chance of a delay?

6. What makes the distribution of diamonds$carat and flights$sched_dep_time similar?

7. Confirm my hypothesis that the early departures of flights in minutes 20-30 and 50-60 are caused by scheduled flights that leave early. Hint: create a binary variable that tells you whether or not a flight was delayed.

## 16.4.5 Time Spans Exercises

1. Why is there months() but no dmonths()?

2. Explain days(overnight * 1) to someone who has just started learning R. How does it work?

3. Create a vector of dates giving the first day of every month in 2015. Create a vector of dates giving the first day of every month in the current year.

4. Write a function that given your birthday (as a date), returns how old you are in years.

5. Why can’t (today() %--% (today() + years(1)) / months(1) work?