---
title: "16 Dates and Times"
author: "Erika Anderson"
date: "September 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load library
library(tidyverse) # basic packages
library(lubridate) # using dates and times
library(nycflights13) # flights data 

# code taken from R Data Science 16.2.2 section
make_datetime_100 <- function(year, month, day, time) {
  make_datetime(year, month, day, time %/% 100, time %% 100)
}

flights_dt <- flights %>% 
  filter(!is.na(dep_time), !is.na(arr_time)) %>% 
  mutate(
    dep_time = make_datetime_100(year, month, day, dep_time),
    arr_time = make_datetime_100(year, month, day, arr_time),
    sched_dep_time = make_datetime_100(year, month, day, sched_dep_time),
    sched_arr_time = make_datetime_100(year, month, day, sched_arr_time)
  ) %>% 
  select(origin, dest, ends_with("delay"), ends_with("time"))


```

A group of DFO coworkers decided to work together to complete the exercises in [R for Data Science](https://r4ds.had.co.nz/) by Garrett Grolemund and Hadley Wickham. 

# 16 Dates and Times

# 16.2.4 Creating Date-Time Exercises

1. What happens if you parse a string that contains invalid dates?

There is an explicit NA.
```{r}
ymd(c("2010-10-10", "bananas"))

```

2. What does the tzone argument to today() do? Why is it important?

tzone assigns a time zone to the today function. This 
```{r}
today() # default to time zone on computer
today(tzone = "UTC") # the date in UTC is not always the same as ours

```


3. Use the appropriate lubridate function to parse each of the following dates:
```{r}
d1 <- "January 1, 2010"
mdy(d1)

d2 <- "2015-Mar-07"
ymd(d2)

d3 <- "06-Jun-2017"
dmy(d3)

d4 <- c("August 19 (2015)", "July 1 (2015)")
mdy(d4)

d5 <- "12/30/14" # Dec 30, 2014
mdy(d5)

```

# Date-time Component Exercises

1. How does the distribution of flight times within a day change over the course of the year?

```{r}

# create column with month as factor
# make all days of month same
# graph departure delays by time of day
flights_dt %>% 
  mutate(dep_hour = update(dep_time, yday = 1),
         month = factor(month(dep_time, label = TRUE))) %>% 
  ggplot(aes(dep_hour, color = month)) +
    geom_freqpoly(binwidth = 300) +
  theme_bw()

# try Jan versus August only
flights_dt %>% 
  mutate(dep_hour = update(dep_time, yday = 1),
         month = factor(month(dep_time, label = TRUE))) %>% 
  filter(month == "Jan" | month == "Aug") %>%
  ggplot(aes(dep_hour, color = month)) +
    geom_freqpoly(binwidth = 300) +
  theme_bw()

# they look similar to me

```


2. Compare dep_time, sched_dep_time and dep_delay. Are they consistent? Explain your findings.

```{r}
# departure delay plus scheduled departure time should be equal to departure time

flights_dt %>%
  mutate(dep_time_calc = sched_dep_time + dep_delay * 60) %>%
  filter(dep_time_calc != dep_time) %>%
  select(dep_time_calc, dep_time, sched_dep_time, dep_delay)

# looks like there are date errors in the calculated departure time
# author talks about them later in the section 16.4 with overnight time spans

```


3. Compare air_time with the duration between the departure and arrival. Explain your findings. (Hint: consider the location of the airport.)

From the flights help file:
dep_time,arr_time  
Actual departure and arrival times (format HHMM or HMM), local tz

This means the air time plus departure time will not be in the same time zone as the arrival time.

```{r}
flights_dt %>%
  mutate(arrival_calc = dep_time + air_time,
         diff = arrival_calc - arr_time) %>%
  filter(arrival_calc != arr_time) %>%
  select(arrival_calc, arr_time, origin, dest, diff) %>%
  arrange(-diff)
```


4. How does the average delay time change over the course of a day? Should you use dep_time or sched_dep_time? Why?

The average delay goes up until 8pm then it goes down again. Likely more flights during the day that causes delays.

```{r}
flights_dt %>%
  mutate(hourOfDay = hour(sched_dep_time)) %>%
  group_by(hourOfDay) %>%
  summarise(meanDepDelay = mean(dep_delay)) %>%

  ggplot(aes(hourOfDay, meanDepDelay)) +
  geom_line()
```



5. On what day of the week should you leave if you want to minimize the chance of a delay?

Saturday

```{r}

flights_dt %>%
  mutate(dayOfWeek = wday(sched_dep_time, label = TRUE)) %>%
  group_by(dayOfWeek) %>%
  summarise(totalDelay = mean(dep_delay, na.rm = TRUE) + mean(arr_delay, na.rm = TRUE)) %>%
  ungroup() %>%

  ggplot(aes(dayOfWeek, totalDelay)) +
  geom_bar(stat = "identity")
```


6. What makes the distribution of diamonds$carat and flights$sched_dep_time similar?

Human friendly peaks for carats and scheduled departure times

```{r}
ggplot(diamonds) +
  geom_density(aes(carat))

ggplot(flights) +
  geom_density(aes(sched_dep_time))
```


7. Confirm my hypothesis that the early departures of flights in minutes 20-30 and 50-60 are caused by scheduled flights that leave early. Hint: create a binary variable that tells you whether or not a flight was delayed.

Looks true for 20 min and 50 min flights. 

```{r}

flights_dt %>%
  mutate(delayed = dep_delay > 0, 
        minute = minute(dep_time)) %>% 
  group_by(minute) %>% 
  summarise(
    avg_delay = mean(arr_delay, na.rm = TRUE),
    n = n()) %>% 
  
  ggplot(aes(minute, avg_delay)) +
    geom_line()

# another way
flights_dt %>%
  mutate(
    minute = minute(dep_time),
    early = dep_delay < 0
  ) %>%
  group_by(minute) %>%
  summarise(
    early = mean(early, na.rm = TRUE),
    n = n()
  ) %>%
  ggplot(aes(minute, early)) +
  geom_line()

```


# Time Span Exercises

1. Why is there months() but no dmonths()?

Duration of a month is variable.

2. Explain days(overnight * 1) to someone who has just started learning R. How does it work?

Overnight is a binary variable, so if the arrival time is before the departure time, it equals one. If it does not, it equals zero. Arrival time then is adjusted by one day, only if the flight is overnight. Otherwise the zero variable drops out of the adjustment.

```{r}

# fix the overnight errors
flights_dt <- flights_dt %>% 
  mutate(
    overnight = arr_time < dep_time,
    arr_time = arr_time + days(overnight * 1),
    sched_arr_time = sched_arr_time + days(overnight * 1)
  )

head(flights_dt)

```


3. Create a vector of dates giving the first day of every month in 2015. Create a vector of dates giving the first day of every month in the current year.

```{r}

# create vector of first day of every month in 2015
# notice it is zero to eleven, 1:12 gives wrong answer
ymd("2015--1-01") + month(0:11)

# create vector of first day of the month for this year
floor_date(today(), unit = "year") + month(0:11)
  

```


4. Write a function that given your birthday (as a date), returns how old you are in years.

```{r}
# this uses integer division 
age <- function(bday) {
  
  # the interval of your birthdate to today, divided by the number of years within that interval
  (bday %--% today()) %/% years(1)
}

age("2001-11-05")
```


5. Why can’t (today() %--% (today() + years(1)) / months(1) work?

Gives 12 months between now and one year from now.Not sure why this is a problem as it gives the right answer. According to Google, this is mixing an interval of dates, with a human period of months. The correct answer uses interval division but it gives the same answer.

```{r}

# this gives today plus one year
today() + years(1)

# this gives interval from today to next year
# removed an extra bracket to work
today() %--% (today() + years(1))

# gives interval of one month
months(1)

# gives the number of months between now and one year from now
(today() %--% (today() + years(1))) / months(1)

# more correct way with integer division
(today() %--% (today() + years(1))) %/% months(1)
```

# Time Zones

Remember to get a list of time zones use

```{r}
head(OlsonNames())
```

