---
title: "19 Functions"
author: "Erika Anderson"
date: "October 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load library
library(tidyverse) # basic packages
library(magrittr) # pipes

```

19.2.1 Practice

1. Why is TRUE not a parameter to rescale01()? What would happen if x contained a single missing value, and na.rm was FALSE?

```{r}

x <- NA

rescale01 <- function(x) {
  rng <- range(x, na.rm = FALSE, finite = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}
rescale01(x)

```

2. In the second variant of rescale01(), infinite values are left unchanged. Rewrite rescale01() so that -Inf is mapped to 0, and Inf is mapped to 1.

```{r}

x <- c(1:10, Inf)

rescale01 <- function(x) {
  rng <- range(x, na.rm = TRUE, finite = TRUE)
  vec <- (x - rng[1]) / (rng[2] - rng[1])
  if_else(vec == -Inf, 0, vec)
  if_else(vec == Inf, 1, vec)
}

rescale01(x)

```

3. Practice turning the following code snippets into functions. Think about what each function does. What would you call it? How many arguments does it need? Can you rewrite it to be more expressive or less duplicative?


```{r}

mean(is.na(x))

meanfn <- function(x) {
  mean(is.na(x))
}

meanfn(x)

x / sum(x, na.rm = TRUE)

thisfn <- function(x) {
  x / sum(x, na.rm = TRUE)
}

thisfn(x)

sd(x, na.rm = TRUE) / mean(x, na.rm = TRUE)

sdfn <- function(x) {
  sd(x, na.rm = TRUE) / meanfn(x)
}

sdfn(x)

```

4. Follow [this link](http://nicercode.github.io/intro/writing-functions.html) to write your own functions to compute the variance and skew of a numeric vector.

5. Write both_na(), a function that takes two vectors of the same length and returns the number of positions that have an NA in both vectors.

```{r}
x <- c(1,2,NA)
y <- c(NA, 2,NA)

both_na <- function(x, y) {
  which(is.na(x) & is.na(y))
}

both_na(x,y)
```

6. What do the following functions do? Why are they useful even though they are so short?

It makes the base R easier to understand what you are doing. It is also easily repeatable.

```{r}

is_directory <- function(x) file.info(x)$isdir
is_readable <- function(x) file.access(x, 4) == 0

```

7. Read the complete lyrics to “Little Bunny Foo Foo”. There’s a lot of duplication in this song. Extend the initial piping example to recreate the complete song, and use functions to reduce the duplication.

```{r}

bunny <- "Little bunny foo foo"

hop <- function(x) {
  "make the bunny hop"
}


```

# 19.3.1 Functions are for Humans Exercises

This function tests whether the given prefix is present in the given string and returns a binomial answer.

```{r}

f1 <- function(string, prefix) {
  substr(string, 1, nchar(prefix)) == prefix
}

f1("test string", "test")

# new name
test_prefix <- function(string, prefix) {
  substr(string, 1, nchar(prefix)) == prefix
}
```

This function returns a null if the input is a single value, and it returns the input minus the last value if the input is greater than a single value.


```{r}
f2 <- function(x) {
  if (length(x) <= 1) return(NULL)
  x[-length(x)]
}

f2(1)
f2(c(1:9))
f2(c("Hello", "Goodbye"))

return_all_except_last <- function(input) {
  if (length(x) <= 1) return(NULL)
  x[-length(x)]
}

```

This function replicates the second input, the same number of values as the first input.

```{r}
f3 <- function(x, y) {
  rep(y, length.out = length(x))
}

f3(1,5)
f3(c(2,5), 8)
f3("Hello", "Goodbye")

replicate_source <- function(amount, source) {
  rep(y, length.out = length(x))
}

```

2. Take a function that you’ve written recently and spend 5 minutes brainstorming a better name for it and its arguments.

I interchanged camel case and snake case in the function name and the data frames. It was hard tor remember what bridge_t_remove was either...transformed data frame with column removed.

```{r}

# make function to create discrete tables for display
breakBridgeTable <- function(df, rowNum, MeasurementNames) {
  
  df <- bridge_t_remove %>%
    select(rowNum) %>%  
    mutate(Tow = MeasurementNames) %>%
    select(Tow, everything())
    
}
```


3. Compare and contrast rnorm() and MASS::mvrnorm(). How could you make them more consistent?

rnorm(n, mean = 0, sd = 1) gives a random normal distribution with given mean and standard deviation.

mvrnorm(n = 1, mu, Sigma, tol = 1e-6, empirical = FALSE, EISPACK = FALSE) produces one or more samples from the specified multivariate normal distribution.

In both there is a "n" parameter, for the number of observations or samples. The mean and mu are the mean(s). I don't see the point of the input called EISPACK since anything other than FALSE produces an error.

Perhaps:

norm_r(n, mean, sd)
norm_mvr(samples, mean, Sigma, tol, empirical)

4. Make a case for why norm_r(), norm_d() etc would be better than rnorm(), dnorm(). Make a case for the opposite.

Using norm as a prefix gives the ability to search related functions using RStudios autocomplete. The current names are easier if you know you want random or density.

# 19.4.4 COnditional Execution Exercises

1. What’s the difference between if and ifelse()? Carefully read the help and construct three examples that illustrate the key differences.

```{r}

x <- 3

if (identical(x, 3)) {message("x is three")} # only has response if correct

ifelse(identical(x, 2), 2, 3) # output has to match the input

if (identical(x, 2)) {message("x is two")} else {message("x is not two")} # gives two possible answers

```


2. Write a greeting function that says “good morning”, “good afternoon”, or “good evening”, depending on the time of day. (Hint: use a time argument that defaults to lubridate::now(). That will make it easier to test your function.)

```{r}

library(lubridate)

greetingfn <- function(time = lubridate::now()) {
  hour <- hour(time)
  
  if (hour < 12) {
    "Good morning"
  } else if (hour < 18) {
    "Good afternoon"
  } else if (hour >= 18) {
    "Good evening"
  } else
    "Hello"
}

greetingfn()
greetingfn(ymd_h("2019-10-23 10"))
greetingfn(ymd_h("2019-10-23 14"))
greetingfn(ymd_h("2019-10-23 18"))
greetingfn(ymd_h("2019-10-23 20"))

```


3. Implement a fizzbuzz function. It takes a single number as input. If the number is divisible by three, it returns “fizz”. If it’s divisible by five it returns “buzz”. If it’s divisible by three and five, it returns “fizzbuzz”. Otherwise, it returns the number. Make sure you first write working code before you create the function.

Use the negative of modulo operator to check divisibility.[What is modular arithmetic?](https://www.khanacademy.org/computing/computer-science/cryptography/modarithmetic/a/what-is-modular-arithmetic)

```{r}

fizzbuzz_fn <- function(x) {

  if (!(x %% 3) && !(x %% 5)) {
    "fizzbuzz"
  } else if (!(x %% 3)) {
    "fizz"
  } else if (!(x %% 5)) {
    "buzz"
  } else {
    x
  }
}

fizzbuzz_fn(6)
fizzbuzz_fn(30)
fizzbuzz_fn(5)
fizzbuzz_fn(11)
fizzbuzz_fn("Hi")


# another way with case_when
fizzbuzz_vec <- function(x) {
  
  # these two lines check that x is a valid input
  stopifnot(length(x) == 1)
  stopifnot(is.numeric(x))
  
  case_when(
    !(x %% 3) & !(x %% 5) ~ "fizzbuzz",
    !(x %% 3) ~ "fizz",
    !(x %% 5) ~ "buzz",
    TRUE ~ as.character(x)
  )
}
fizzbuzz_vec(c(0, 1, 2, 3, 5, 9, 10, 12, 15))

```


4. How could you use cut() to simplify this set of nested if-else statements?

```{r}

# simple data
temp <- 5

# given this
if (temp <= 0) {
  "freezing"
} else if (temp <= 10) {
  "cold"
} else if (temp <= 20) {
  "cool"
} else if (temp <= 30) {
  "warm"
} else {
  "hot"
}

# example data
temp <- c(-5, 5, 15, 25, 35)

# use cut function to simplify
cut(temp, 
    breaks = c(-Inf, 0, 10, 20, 30, Inf),
    labels = c("freezing", "cold", "cool", "warm", "hot"))

```


5. How would you change the call to cut() if I’d used < instead of <=? What is the other chief advantage of cut() for this problem? (Hint: what happens if you have many values in temp?)

The option right = TRUE would make the cut function < instead of <=
The main advantage is that cut works on vectors, whereas if works only on a single value.


6. What happens if you use switch() with numeric values?

You are supposed to use integer values with switch. If you use numeric values, the function ignores everything past the decimal. It does not round the numeric, just cuts it off.

```{r}

# test switch with integer values
switch(1, 2, 3, 4, 5)
switch(3, "Hi", "Bye", "Oh")

# test switch with numeric values
switch(3.5, 2, 3, 4, 5)
switch(2.7, "Hi", "Bye", "Oh")
```


7. What does this switch() call do? What happens if x is “e”?

Nothing happens with x is "e" or the integer is 5.The function returns NULL

```{r}

# gives error with x <- 1
# Error: empty alternative in numeric switch
x <- 2

switch(x, 
  a = ,
  b = "ab",
  c = ,
  d = "cd"
)

x <- "e"

switch(x, 
  a = ,
  b = "ab",
  c = ,
  d = "cd"
)

```

19.5.5 Function Arguments Exercises

1. What does commas(letters, collapse = "-") do? Why?

2. It’d be nice if you could supply multiple characters to the pad argument, e.g. rule("Title", pad = "-+"). Why doesn’t this currently work? How could you fix it?

3. What does the trim argument to mean() do? When might you use it?

4. The default value for the method argument to cor() is c("pearson", "kendall", "spearman"). What does that mean? What value is used by default?

