---
title: "21 Iteration"
date: "November 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load library
library(tidyverse) # basic packages
library(magrittr) # pipes
library(purrr) # functional programming ie map

```

# 21.1.1 For Loop Exercises

1. Write for loops to:

-Compute the mean of every column in mtcars.
```{r}

# create empty vector
output <- vector("double", ncol(mtcars))

# make named vector
names(output) <- names(mtcars)

# loop
for (i in seq_along(mtcars)) {
  output[[i]] <- mean(mtcars[[i]], na.rm = TRUE)
}

output
```


-Determine the type of each column in nycflights13::flights.
```{r}

library(nycflights13) # get data

# get names of columns
flightColumns <- names(flights)

# create empty list
### notice that list is needed rather than a vector since the time_hour has two values from class ###
output <- vector("list", length(flightColumns))

# name vector
names(output) <- flightColumns

for (i in seq_along(flightColumns)) {
output[[i]] <- class(flights[[i]])
  
}

output
```

-Compute the number of unique values in each column of iris.

```{r}

# create an empty list
output <- vector("list", ncol(iris))

# name the output 
names(output) <- names(iris)

for (i in seq_along(names(iris))) {
  output[[i]] <- length(unique(iris[[i]]))
  
}

output

```

-Generate 10 random normals for each (-10, 0, 10, 100).Think about the output, sequence, and body before you start writing the loop.

```{r}

# given vector
x <- c(-10, 0, 10, 100)

# create empty output
output <- vector("list", length(x))

# sequence
for (i in seq_along(x)) {
  output[[i]] <- rnorm(10, i)
}

output

```


2. Eliminate the for loop in each of the following examples by taking advantage of an existing function that works with vectors:

```{r}

out <- ""
for (x in letters) {
  out <- stringr::str_c(out, x)
}

out

# redo without loop
out2 <- paste0(letters, collapse = "")
out2

# redo without loop
out3 <- str_c(letters, collapse = "")
out3

```


```{r}
x <- sample(100)
sd <- 0
for (i in seq_along(x)) {
  sd <- sd + (x[i] - mean(x)) ^ 2
}
sd <- sqrt(sd / (length(x) - 1))
sd

# redo without loop using standard deviation function
sd(x)

```


```{r}
x <- runif(100)
out <- vector("numeric", length(x))
out[1] <- x[1]
for (i in 2:length(x)) {
  out[i] <- out[i - 1] + x[i]
}
out

# without loop find cumulative sum
cumsum(x)
```

3. Combine your function writing and for loop skills:

-Write a for loop that prints() the lyrics to the children’s song “Alice the camel”.
```{r}

x <- c("three", "two", "one")

for (i in seq_along(x)) {
  line <- str_c("Alice the camel has ", rep(x[i],3), " humps.\n", "\n", collapse = "\n")
  verse <- str_c(line, " So Go Alice Go\n")
  cat(verse)
}

```


-Convert the nursery rhyme “ten in the bed” to a function. Generalise it to any number of people in any sleeping structure.
```{r}

output <- vector("character", 3)

x <- c("three", "two", "one")
y <- head(x, -1)

for (i in seq_along(y)) {
  
  output[[i]] <- paste("There were", x[i], "in the bed and the little one said, Roll Over, roll over. So they all rolled over and one fell out. \n")
}
  
# create last verse
 finale <- "There was one in the bed and the little one said, I'm lonely...\n"

# print song
cat(output, finale)



```


-Convert the song “99 bottles of beer on the wall” to a function. Generalise to any number of any vessel containing any liquid on any surface.
```{r}

# create function
beerSong <- function(numBottles, liquid, surface) {
  
  # should have tests for input variable here with stopifnot with wrong type
  
  # make a descending vector of number of bottles
  x <- numBottles:0
  
  # create empty vector for output
  output <- vector("character", numBottles + 1)
  
  # sequence along numbers
  for (i in seq_along(numBottles)) {
    
    # create number of bottles in text form
    j <- x[[i]]
    
    if (j == 0) {
      numBottles <- "No more bottles"
    } else if (j == 1) {
      numBottles <- "1 bottle"
    } else {
      numBottles <- str_c(j, " bottles")
    }
    
    
    # need to modify to have one less bottle in last part of sentence
    verse <- str_c(numBottles, " of ", liquid, " on the ", surface, ", ", numBottles, " of ", liquid, ". Take one down and pass it around. ", numBottles, " of ", liquid, ".")
    
    output[[i]] <- verse
  }
  
  # add last line to verses
  output[[i + 1]] <- str_c("Let's go buy more ", liquid, "!")
  
  cat(output)
}

# test function
beerSong(3, "beer", "wall")

# test with other parameters
#beerSong(2, "milk", "table")

```


4. It’s common to see for loops that don’t preallocate the output and instead increase the length of a vector at each step:


```{r}

output <- vector("integer", 0)
for (i in seq_along(x)) {
  output <- c(output, lengths(x[[i]]))
}
output

```
How does this affect performance? Design and execute an experiment.

It doubles the time in simple functions if you incrementally add to a vector. The morale is use a premade empty vector for output.

```{r}

library(microbenchmark) # used to compare speeds of functions

# given loop that adds to vector
# make function

addFunction <- function(x) {
  output <- vector("integer", 0)
  for (i in seq_along(x)) {
    output <- c(output, lengths(x[[i]]))
  }
}

# modified function that uses premade empty vector
premadeVectorFn <- function(x) {
  output2 <- vector("integer", length(x))
  for (i in seq_along(x)) {
    output2[[i]] <- 1
  }
}

# compare speed of functions
microbenchmark(addFunction(1000), premadeVectorFn(1000), times = 10)

```


21.3.5 For loop variations Exercises

1. Imagine you have a directory full of CSV files that you want to read in. You have their paths in a vector, , and now want to read each one with read_csv(). Write the for loop that will load them into a single data frame.

```{r}

# list of files
files <- dir("data/", pattern = "\\.csv$", full.names = TRUE)

# create list
output <- vector("list", length(files))

for (i in seq_along(files)) { 
  x <- read_csv(files[[i]],
                col_names = TRUE,
                cols(
                  A = col_double(),
                  B = col_double(),
                  C = col_double()
                ))
  output[[i]] <- x
}

# flatten into one dataframe
bind_rows(output)

```


2. What happens if you use for (nm in names(x)) and x has no names? What if only some of the elements are named? What if the names are not unique?

-Nothing happens, and no error is given, if you loop using names on an unamed vector
-Using loop with some elements with names gives an error: Error in x[[nm]] : subscript out of bounds
-Works (but selects only the first named duplicate element without error), but if you use the name to pull the value out there is uses the first element only both times

```{r}
# create unnamed vector
x <- 1:3

for (nm in names(x)) {
  cat(nm)
}

# create vector with some elements named
x <- c("Hi" = 1, "Bye" = 2, 3)


# for (nm in names(x)) {
#   cat(nm)
#   cat(x[[nm]])
# }


# create named vector with duplicate name
x <- 1:3
names(x) <- c("Hi", "Bye", "Hi")

for (nm in names(x)) {
  cat(x[[nm]])
}

```


3. Write a function that prints the mean of each numeric column in a data frame, along with its name. For example, show_mean(iris) would print:
Sepal.Length: 5.84  
Sepal.Width:  3.06  
Petal.Length: 3.76  
Petal.Width:  1.20  

```{r}

# create function
show_mean <- function(df) {
  
# vector to sequence along
vector <- 1:ncol(df)
names(vector) <- colnames(df)

# loop through each column
for (i in vector) {
  
  column <- df[ ,i]
  
  if (is.numeric(column) == TRUE) {
    
    cat(
      str_c(
        names(vector)[vector == i]),
        ":  ",
        round(mean(column),2),
        "\n")
  } 
}

}

# test function
show_mean(iris)
```

(Extra challenge: what function did I use to make sure that the numbers lined up nicely, even though the variable names had different lengths?)

4. What does this code do? How does it work?

```{r}

# this creates a named list of two functions
trans <- list( 
  
  # this function called disp is multiplied by a constant
  # cubic inch to cubic meter
  disp = function(x) x * 0.0163871,
  
  # this function creates a function that creates a factor with two levels called auto and manual
  am = function(x) {
    factor(x, labels = c("auto", "manual"))
  }
)

# this sequences along the variables named in the list using the same functions within the trans list
for (var in names(trans)) {
  
  # this overwrites the mtcars data frame using the functions in the list
  mtcars[[var]] <- trans[[var]](mtcars[[var]])
}
```

21.4.1 For loop vs. functional Exercises

1. Read the documentation for apply(). In the 2d case, what two for loops does it generalise?

apply either loops over columns or rows determined by the second MARGIN argument of 1 for rows or 2 for columns. You can use c(1,2) if you want to loop over both. jrnold says these represent seq_len(nrow(x)) and seq_len(ncol(x)) but I am not sure this works with the 2d case. Any ideas? 

2. Adapt col_summary() so that it only applies to numeric columns You might want to start with an is_numeric() function that returns a logical vector that has a TRUE corresponding to each numeric column.

```{r}

# original data frame
df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)

# original function given
col_summary <- function(df, fun) {
  out <- vector("double", length(df))
  for (i in seq_along(df)) {
    out[i] <- fun(df[[i]])
  }
  out
}

# test function with median
col_summary(df, median)

# test function with mean
col_summary(df, mean)

# create data frame with numeric and non-numeric columns
df2 <- tibble(
  A = 1:3,
  B = c("Hi", "Bye", "Yes"),
  C = 4:6
)

# create new function that only works on numeric columns
col_summary2 <- function(df2, fn) {
  
# create empty vector to hold results of numeric test
numericOut <- vector("logical", ncol(df2))

# name the vector with column names
names(numericOut) <- names(df2)


for (i in seq_along(df2)) {
  numericOut[[i]] <- is.numeric(df2[[i]])
}

# find which columns are numeric
indNumeric <- which(numericOut == TRUE)
names(indNumeric) <- names(which(numericOut == TRUE))

# create empty vector for results of function applied to only numeric columns
out <- vector("double", length(indNumeric))
names(out) <- names(indNumeric)

  for (i in seq_along(indNumeric)) {
    x <- indNumeric[i]
    out[i] <- fn(df2[[x]])
  }
  out
}

# test new function
col_summary2(df2, mean)

```

21.5.3 Map Function Exercises

1. Write code that uses one of the map functions to:

-Compute the mean of every column in mtcars.

```{r}

library(tidyverse)

map_dbl(mtcars, mean)

mtcars %>%
  map_dbl(mean)

```

-Determine the type of each column in nycflights13::flights.

```{r}

nycflights13::flights %>%
map_chr(typeof)

```

-Compute the number of unique values in each column of iris.

```{r}

# descriptive way
map_int(iris, function(x) length(unique(x)))

# using shorter function methods
map_int(iris, ~length(unique(.)))

# using pipes
iris %>%
  map_int(~length(unique(.)))

```

-Generate 10 random normals for each of  −10, 0, 10, and 100.

```{r}

# create data set
x <- c(-10, 0, 10, 100)

# standard way
map(x, ~rnorm(10, mean = .))

# with pipes
x %>%
  map(~rnorm(10, mean = .))

```


2. How can you create a single vector that for each column in a data frame indicates whether or not it’s a factor?

```{r}

str(iris)

map_lgl(iris, is.factor)

```


3. What happens when you use the map functions on vectors that aren’t lists? What does map(1:5, runif) do? Why?

If the input is a vector, map pretends each value is a portion of a list because map expects a list as input. 
```{r}

map(1:5, runif)

```

4. What does map(x, ~rnorm, n = 5) do? Why? What does map_dbl(x, ~rnorm, n = 5) do with x a vector from -2 to 2? Why?

```{r}

# this gives five random normals for -2 to 2
map(-2:2, rnorm, n = 5)

# this gives an error because the result is a list, and not the expected double vector
# map_dbl(-2:2, rnorm, n = 5) 

```

5. Rewrite map(x, function(df) lm(mpg ~ wt, data = df)) to eliminate the anonymous function.

```{r}

# make data set to test
x <- mtcars %>%
  split(.$cyl)

cat("Given function\n")
map(x, function(df) lm(mpg ~ wt, data = df))

cat("Shorter Version\n")
map(x, ~lm(mpg ~ wt, data = .))

cat("With pipes\n")
x %>%
  map(~lm(mpg ~ wt, data = .))

```

