---
title: "23 Model Basics"
date: "January 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load library
library(tidyverse) # basic packages
library(magrittr) # pipes
library(modelr) # modelling

# suggested option
options(na.action = na.warn)

```

# 23.1.1 A Simple Model Exercises

1. One downside of the linear model is that it is sensitive to unusual values because the distance incorporates a squared term. Fit a linear model to the simulated data below, and visualise the results. Rerun a few times to generate different simulated datasets. What do you notice about the model?

```{r}

sim1a <- tibble(
  x = rep(1:10, each = 3),
  y = x * 1.5 + 6 + rt(length(x), df = 2)
)

```

2. One way to make linear models more robust is to use a different distance measure. For example, instead of root-mean-squared distance, you could use mean-absolute distance:

```{r}

measure_distance <- function(mod, data) {
  diff <- data$y - model1(mod, data)
  mean(abs(diff))
}

```

Use optim() to fit this model to the simulated data above and compare it to the linear model.

3. ne challenge with performing numerical optimisation is that it’s only guaranteed to find one local optimum. What’s the problem with optimising a three parameter model like this?

```{r}

model1 <- function(a, data) {
  a[1] + data$x * a[2] + a[3]
}

```
# 23.3.3 Visualising Models Exercises

1. Instead of using lm() to fit a straight line, you can use loess() to fit a smooth curve. Repeat the process of model fitting, grid generation, predictions, and visualisation on sim1 using loess() instead of lm(). How does the result compare to geom_smooth()?

2. add_predictions() is paired with gather_predictions() and spread_predictions(). How do these three functions differ?

3. What does geom_ref_line() do? What package does it come from? Why is displaying a reference line in plots showing residuals useful and important?

4.Why might you want to look at a frequency polygon of absolute residuals? What are the pros and cons compared to looking at the raw residuals?



