---
title: "24 Model Building"
date: "March 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load libraries
library(tidyverse)
library(modelr)
options(na.action = na.warn)

library(nycflights13)
library(lubridate)

```

# 24.2.3 Why are low quality diamonds more expensive? Exercises

1. In the plot of lcarat vs. lprice, there are some bright vertical strips. What do they represent?

These vertical strips represent the even numbers associated with the size of diamond to market to people.


```{r}

# given
diamonds2 <- diamonds %>% 
  filter(carat <= 2.5) %>% 
  mutate(lprice = log2(price), 
         lcarat = log2(carat))

ggplot(diamonds2, aes(lcarat, lprice)) + 
  geom_hex(bins = 50)


```


2. If log(price) = a_0 + a_1 * log(carat), what does that say about the relationship between price and carat?

The log relationship means there is an exponential relationship between carat and price, rather than linear. So if the carat goes up by 2, then the price goes up by a power. See jarnold for more details.

```{r}

# given
mod_diamond <- lm(lprice ~ lcarat, data = diamonds2)

grid <- diamonds2 %>% 
  data_grid(carat = seq_range(carat, 20)) %>% 
  mutate(lcarat = log2(carat)) %>% 
  add_predictions(mod_diamond, "lprice") %>% 
  mutate(price = 2 ^ lprice)

ggplot(diamonds2, aes(carat, price)) + 
  geom_hex(bins = 50) + 
  geom_line(data = grid, colour = "red", size = 1)

```


3. Extract the diamonds that have very high and very low residuals. Is there anything unusual about these diamonds? Are they particularly bad or good, or do you think these are pricing errors?

More smaller carat, fair cut, and F color diamonds in the residuals. I suspect that it is overpricing diamonds.

```{r}

# given
mod_diamond2 <- lm(lprice ~ lcarat + color + cut + clarity, data = diamonds2)

grid <- diamonds2 %>% 
  data_grid(cut, .model = mod_diamond2) %>% 
  add_predictions(mod_diamond2)

diamonds2 <- diamonds2 %>% 
  add_residuals(mod_diamond2, "lresid2")

# select outliers above and below 1, -1 residuals
diamonds3 <- diamonds2 %>%
  filter(lresid2 > 1 | lresid2 < -1)
diamonds3

# plot them
ggplot(diamonds3, aes(carat, lresid2)) + 
  geom_point()

ggplot(diamonds3, aes(color, lresid2)) + 
  geom_point()

ggplot(diamonds3, aes(cut, lresid2)) + 
  geom_point()


ggplot(diamonds3, aes(clarity, lresid2)) + 
  geom_point()

```



4. Does the final model, mod_diamond2, do a good job of predicting diamond prices? Would you trust it to tell you how much to spend if you were buying a diamond?

Yes you could use this model to tell you where the price fits into the overall picture. To compare to other models, you could compare Root mean sum squares or the mean absolute residual. Or I may use AIC.

It pays to use common sense, if it looks too low, there may be another reason that you haven/t considered. Is it stolen? If it is higher, perhaps there is a historic reason the diamond is valued.

```{r}
# given
mod_diamond2 <- lm(lprice ~ lcarat + color + cut + clarity, data = diamonds2)

# untransform the prediction to put price in useful value
grid <- diamonds2 %>% 
  data_grid(cut, .model = mod_diamond2) %>% 
  add_predictions(mod_diamond2) %>%
  mutate(PriceTrans = 2^pred)

# plot the cut by price to see the value
ggplot(grid, aes(cut, PriceTrans)) + 
  geom_point()

# could do this for all the other factors

# janorld recommends calculating some stats on the residuals to quantify the model
lresid2_summary <- summarise(diamonds2,
  rmse = sqrt(mean(lresid2^2)),
  mae = mean(abs(lresid2)),
  p025 = quantile(lresid2, 0.025),
  p975 = quantile(lresid2, 0.975)
)
lresid2_summary

```

# 24.3.5 Flights and Days of Week Exercises

1. Use your Google sleuthing skills to brainstorm why there were fewer than expected flights on Jan 20, May 26, and Sep 1. (Hint: they all have the same explanation.) How would these days generalise to another year?

2. What do the three days with high positive residuals represent? How would these days generalise to another year?

```{r}
daily %>% 
  top_n(3, resid)

```

3. Create a new variable that splits the wday variable into terms, but only for Saturdays, i.e. it should have Thurs, Fri, but Sat-summer, Sat-spring, Sat-fall. How does this model compare with the model with every combination of wday and term?

. Create a new wday variable that combines the day of week, term (for Saturdays), and public holidays. What do the residuals of that model look like?

5. What happens if you fit a day of week effect that varies by month (i.e. n ~ wday * month)? Why is this not very helpful?

6. What would you expect the model n ~ wday + ns(date, 5) to look like? Knowing what you know about the data, why would you expect it to be not particularly effective?

7. We hypothesised that people leaving on Sundays are more likely to be business travellers who need to be somewhere on Monday. Explore that hypothesis by seeing how it breaks down based on distance and time: if it’s true, you’d expect to see more Sunday evening flights to places that are far away.

8. It’s a little frustrating that Sunday and Saturday are on separate ends of the plot. Write a small function to set the levels of the factor so that the week starts on Monday.

