---
title: "25 Many Model"
date: "April 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load libraries
library(modelr)
library(tidyverse)
library(gapminder) # dataset on life expectancy and GDP
library(broom) # functions for tidy data with models
library(ggbeeswarm) # other methods to jitter

```

25.2.5 gapminder Exercises

1. A linear trend seems to be slightly too simple for the overall trend. Can you do better with a quadratic polynomial? How can you interpret the coefficients of the quadratic? (Hint you might want to transform year so that it has mean zero.)

A quadratic polynomial is the format life expectancy = coefficient * (year)2 + residual error.
The coeficients gives the direction of change (+ or -), and the amount the life expectancy would change per year. If the year is zero, the coefficient drops out and the remaining term is the average expected life span of a person in that country. For example Canada is 74.9 years over 1952-2007. 

The quadratic polynomial fits the African curves better than the linear model, improving their R2 values. Nevertheless, Rwanda still is an outlier with a low r2 value even with the new model. The dip is due to a civil war in 1994, when there was a mass slaughter of Tutsi. Up to one million people were killed.

```{r}

# using given code
by_country <- gapminder %>% 
  group_by(country, continent) %>% 
  nest()

# new quadratic polynomial model 
# using the hint by subtracting the median year within the function
country_model_quad <- function(df) {
  lm(lifeExp ~ poly(year - median(year), 2), data = df)
}

# apply new model to data frame
by_country <- by_country %>% 
  mutate(model = map(data, country_model_quad),
  
  # add residuals
  resids = map2(data, model, add_residuals))

resids <- unnest(by_country, resids)

resids %>% 
  ggplot(aes(year, resid, group = country)) +
    geom_line(alpha = 1 / 3) + 
    facet_wrap(~continent)

glance <- by_country %>% 
  mutate(glance = map(model, broom::glance)) %>% 
  unnest(glance, .drop = TRUE)

glance %>% 
  ggplot(aes(continent, r.squared)) + 
    geom_jitter(width = 0.5)

bad_fit <- filter(glance, r.squared < 0.25)

gapminder %>% 
  semi_join(bad_fit, by = "country") %>% 
  ggplot(aes(year, lifeExp, colour = country)) +
    geom_line()

# find out coeffiecents for Canada
Canada <- by_country %>%
  filter(country == "Canada")
  
Canada$model

```


2. Explore other methods for visualising the distribution of r2 per continent. You might want to try the ggbeeswarm package, which provides similar methods for avoiding overlaps as jitter, but uses deterministic methods.

```{r}

# violin scatterplot
glance %>% 
  ggplot(aes(continent, r.squared)) + 
    geom_quasirandom()

# another option
glance %>% 
  ggplot(aes(continent, r.squared)) + 
    geom_beeswarm(aes(color = continent), dodge.width=.8)

# use facets within beeswarm and AIC
glance %>%
  ggplot(aes(AIC, r.squared)) +
  geom_quasirandom(groupOnX = FALSE) +
  facet_grid(.~continent)

```


3. To create the last plot (showing the data for the countries with the worst model fits), we needed two steps: we created a data frame with one row per country and then semi-joined it to the original dataset. It’s possible to avoid this join if we use unnest() instead of unnest(.drop = TRUE). How?

Use unnest on individual columns to pull out details.

```{r}

by_country %>% 
  mutate(glance = map(model, broom::glance)) %>% 
  #unnest(cols = c(glance, data)) %>%
  unnest(glance) %>%
  unnest(data) %>%
  filter(r.squared < 0.25) %>%

  ggplot(aes(year, lifeExp, colour = country)) +
    geom_line()

```


25.4.5 Creating List Columns Exercises

1. List all the functions that you can think of that take a atomic vector and return a list.

Many of the splitting functions with stringr and stringi give lists. A list column would be useful when different numbers of results are returned for different rows. Many models return lists, and spatial functions are often returned as lists. Which is why this section is so useful, if mind bending.


```{r}
# make data frame
dat <- tibble(a, b = (a * 2))

# make model
mod <- lm(b ~ a, data = dat)
mod
str(mod)

```


2. Brainstorm useful summary functions that, like quantile(), return multiple values.

Jrnold duggests range and boxplot.stats. I like the boxplot.stats and will consider using it. Generally, I have used mutate with min and max to get range within dataframe, but see the simplity to using range with a list column.


3. What’s missing in the following data frame? How does quantile() return that missing piece? Why isn’t that helpful here?

This is a repeat from the text, and prob for each quantile is missing after unnest. But there is a way to include it.

```{r}
# code given
mtcars %>% 
  group_by(cyl) %>% 
  summarise(q = list(quantile(mpg))) %>% 
  unnest(cols = c(q))

# which quantiles are given?
?quantile
seq(0, 1, 0.25)

# another way to get quantiles and their prob
# I changed it to use the default quantiles so I could compare the data frames
probs <- seq(0, 1, 0.25)
mtcars %>% 
  group_by(cyl) %>% 
  summarise(p = list(probs), q = list(quantile(mpg, probs))) %>% 
  unnest(cols = c(p, q))

```

4, What does this code do? Why might might it be useful?

This code groups all the rows from mtcars by cylinder, creating list columns for each other parameter with a list of all the values for that number of cylinders. It may be useful is you want to apply a function to each group of cylinders using the original values of parameters. I rewrote with newer version of code given the warning that funs() is deprecated now (seee below). 


```{r}
given4 <- mtcars %>% 
  group_by(cyl) %>% 
  summarise_each(funs(list))
given4

# let's look at what's inside
cat("mpg\n")
given4$mpg
cat("\nhp\n")
given4$hp
# note that this gives index values and not the number of cylinders

# can we rewrite with newer version given in warning?
mtcars %>%
  group_by(cyl) %>%
  summarise_each(lst(list)) 
# where does the second list come from? Is is just a function? 
# then why two lst(list) needed


```

Copy of Warning Given only once per session: 

funs() is soft deprecated as of dplyr 0.8.0
Please use a list of either functions or lambdas: 

  # Simple named list: 
  list(mean = mean, median = median)

  # Auto named with `tibble::lst()`: 
  tibble::lst(mean, median)

  # Using lambdas
  list(~ mean(., trim = .2), ~ median(., na.rm = TRUE))
This warning is displayed once per session.


# 25.5.3 Simplifiying List Columns Exercises

1. Why might the lengths() function be useful for creating atomic vector columns from list-columns?

If lengths is the same, then unnest could work. So length function could be used to test if unnest would work.

2. List the most common types of vector found in a data frame. What makes lists different?

character, double, integer, logical
Lists are different because they can contain more than one type of value and difference amounts.

# Making tidy data with broom

The functions glance (gives model stats), tidy (gives model cooeffiecents) and augment (adds predictions, residuals, and se to data frame) are useful.

