---
title: "25 Many Model"
date: "April 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load libraries
library(modelr)
library(tidyverse)
library(gapminder) # dataset on life expectancy and GDP
library(broom) # functions for tidy data with models
library(ggbeeswarm) # other methods to jitter

```

25.2.5 gapminder Exercises

1. A linear trend seems to be slightly too simple for the overall trend. Can you do better with a quadratic polynomial? How can you interpret the coefficients of the quadratic? (Hint you might want to transform year so that it has mean zero.)

A quadratic polynomial is the format life expectancy = coefficient * (year)2 + residual error.
The coeficients gives the direction of change (+ or -), and the amount the life expectancy would change per year. If the year is zero, the coefficient drops out and the remaining term is the average expected life span of a person in that country. For example Canada is 74.9 years over 1952-2007. 

The quadratic polynomial fits the African curves better than the linear model, improving their R2 values. Nevertheless, Rwanda still is an outlier with a low r2 value even with the new model. The dip is due to a civil war in 1994, when there was a mass slaughter of Tutsi. Up to one million people were killed.

```{r}

# using given code
by_country <- gapminder %>% 
  group_by(country, continent) %>% 
  nest()

# new quadratic polynomial model 
# using the hint by subtracting the median year within the function
country_model_quad <- function(df) {
  lm(lifeExp ~ poly(year - median(year), 2), data = df)
}

# apply new model to data frame
by_country <- by_country %>% 
  mutate(model = map(data, country_model_quad),
  
  # add residuals
  resids = map2(data, model, add_residuals))

resids <- unnest(by_country, resids)

resids %>% 
  ggplot(aes(year, resid, group = country)) +
    geom_line(alpha = 1 / 3) + 
    facet_wrap(~continent)

glance <- by_country %>% 
  mutate(glance = map(model, broom::glance)) %>% 
  unnest(glance, .drop = TRUE)

glance %>% 
  ggplot(aes(continent, r.squared)) + 
    geom_jitter(width = 0.5)

bad_fit <- filter(glance, r.squared < 0.25)

gapminder %>% 
  semi_join(bad_fit, by = "country") %>% 
  ggplot(aes(year, lifeExp, colour = country)) +
    geom_line()

# find out coeffiecents for Canada
Canada <- by_country %>%
  filter(country == "Canada")
  
Canada$model

```


2. Explore other methods for visualising the distribution of r2 per continent. You might want to try the ggbeeswarm package, which provides similar methods for avoiding overlaps as jitter, but uses deterministic methods.

```{r}

# violin scatterplot
glance %>% 
  ggplot(aes(continent, r.squared)) + 
    geom_quasirandom()

# another option
glance %>% 
  ggplot(aes(continent, r.squared)) + 
    geom_beeswarm(aes(color = continent), dodge.width=.8)

# use facets within beeswarm and AIC
glance %>%
  ggplot(aes(AIC, r.squared)) +
  geom_quasirandom(groupOnX = FALSE) +
  facet_grid(.~continent)

```


3. To create the last plot (showing the data for the countries with the worst model fits), we needed two steps: we created a data frame with one row per country and then semi-joined it to the original dataset. It’s possible to avoid this join if we use unnest() instead of unnest(.drop = TRUE). How?

Use unnest on individual columns to pull out details.

```{r}

by_country %>% 
  mutate(glance = map(model, broom::glance)) %>% 
  #unnest(cols = c(glance, data)) %>%
  unnest(glance) %>%
  unnest(data) %>%
  filter(r.squared < 0.25) %>%

  ggplot(aes(year, lifeExp, colour = country)) +
    geom_line()

```


25.4.5 Creating List Columns Exercises

1. List all the functions that you can think of that take a atomic vector and return a list.

2. Brainstorm useful summary functions that, like quantile(), return multiple values.

3. What’s missing in the following data frame? How does quantile() return that missing piece? Why isn’t that helpful here?

```{r}
# given
mtcars %>% 
  group_by(cyl) %>% 
  summarise(q = list(quantile(mpg))) %>% 
  unnest()

```

4, What does this code do? Why might might it be useful?

```{r}
mtcars %>% 
  group_by(cyl) %>% 
  summarise_each(funs(list))

```

