---
title: "27 R Markdown"
date: "May 2020"
output: 
      html_document: default
      pdf_document: default

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load libraries
library(tidyverse)

```

## 27.2.1 R Markdown Basicis Exercises

1. Create a new notebook using File > New File > R Notebook. Read the instructions. Practice running the chunks. Verify that you can modify the code, re-run it, and see modified output.

Do togther

2. Create a new R Markdown document with File > New File > R Markdown… Knit it by clicking the appropriate button. Knit it by using the appropriate keyboard short cut. Verify that you can modify the input and see the output update.

Keyboard shortcut is Cntrl + Shift + K


3. Compare and contrast the R notebook and R markdown files you created above. How are the outputs similar? How are they different? How are the inputs similar? How are they different? What happens if you copy the YAML header from one to the other?

The biggest difference is that the R code is not run during the knit in the notebook. If you change the header to the same as the R markdown file, the notebook becomes a R markdown file and runs the R code when it knits. 

You can also use the knitted the notebook with the nb.html file to reproduce the R notebook, but you cannot use the html or pdf to reproduce the R markdown file

In the html file you can interactively hide and show the code.

4. Create one new R Markdown document for each of the three built-in formats: HTML, PDF and Word. Knit each of the three documents. How does the output differ? How does the input differ? (You may need to install LaTeX in order to build the PDF output — RStudio will prompt you if this is necessary.)

You need to change the header to word_document or pdf_document, or use the drop down arrow beside knit to make the different formats. It is also good to know that you can produce more than one format with the file by changing the header output yaml to output: pdf_document: default word_document:: default. 

```{r}
# another way to render multiple formats using R code in console
#rmarkdown::render(<your-rmd-file.rmd>, output_format ="all")
```


Generally, html is preferred to work in as it can become any other format but some of the others cannot always go backwards.

## 27.3.1 Exercises

1. Practice what you’ve learned by creating a brief CV. The title should be your name, and you should include headings for (at least) education or employment. Each of the sections should include a bulleted list of jobs/degrees. Highlight the year in bold.

see PracticeCV.Rmd file

2. Using the R Markdown quick reference, figure out how to:

  * Add a footnote.
^[this is my footnote]

  
  * Add a horizontal rule.
    three or more astericks or dashes
    
****

____
  
  * Add a block quote.
  
> Use the greater than symbol to add block quotes
  
3. Copy and paste the contents of diamond-sizes.Rmd from <https://github.com/hadley/r4ds/tree/master/rmarkdown> in to a local R markdown document. Check that you can run it, then add text after the frequency polygon that describes its most striking features.

[1]: <https://r4ds.had.co.nz/r-markdown.html#exercises-71> 

## 27.4.7 Code Chunks Exercises

see DiamondSizes.Rmd for all four exercises

1. Add a section that explores how diamond sizes vary by cut, colour, and clarity. Assume you’re writing a report for someone who doesn’t know R, and instead of setting echo = FALSE on each chunk, set a global option.

2. Download diamond-sizes.Rmd from https://github.com/hadley/r4ds/tree/master/rmarkdown. Add a section that describes the largest 20 diamonds, including a table that displays their most important attributes.

3. Modify diamonds-sizes.Rmd to use comma() to produce nicely formatted output. Also include the percentage of diamonds that are larger than 2.5 carats.

4. Set up a network of chunks where d depends on c and b, and both b and c depend on a. Have each chunk print lubridate::now(), set cache = TRUE, then verify your understanding of caching.

