---
title: "28 Graphics for Communication"
date: "June 2020"
output: 
      html_document: default
      
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load libraries
library(tidyverse)

# change default graph type to one I like
theme_set(theme_bw())

# load libraries for my graphic
library(ggmap) # map of canada
library(ggpubr) # panel of plots
library(rgdal) # shapefiles


```

## 28.2.1 Label Exercises

1.Create one plot on the fuel economy data with customised title, subtitle, caption, x, y, and colour labels.

```{r}
# given but change labels
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  labs(title = "Fuel economy changes by car type",
       subtitle = "(Two seater sports cars are the exception)",
       caption = "Exercise 28.2.1",
    x = "Engine displacement (L)",
    y = "Highway fuel economy (mpg)",
    colour = "Car type"
  )

```


2.The geom_smooth() is somewhat misleading because the hwy for large engines is skewed upwards due to the inclusion of lightweight sports cars with big engines. Use your modelling tools to fit and display a better model.

```{r}

ggplot(mpg, aes(displ, hwy, color = class)) +
  geom_point(aes(color = class)) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Different fuel effciency in different car classes",
       subtitle = "All negative relationships except 2 seaters",
       caption = "Exercise 28.2.2",
    x = "Engine displacement (L)",
    y = "Highway fuel economy (mpg)",
    colour = "Car type"
  )

# try different models
mod_displ <- lm(hwy ~ displ, data = mpg)
mod_class <- lm(hwy ~ class, data = mpg)
mod_displ_class <- lm(hwy ~ displ + class, data = mpg)
mod_displ_class_interact <- lm(hwy ~ displ * class, data = mpg)

# compare models and choose smallest AIC
AIC(mod_displ, mod_class, mod_displ_class, mod_displ_class_interact)

# graph residuals
mpg %>% 
  modelr::add_residuals(., mod_displ_class_interact) %>%
  ggplot(aes(displ, resid, color = class)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(title = "Car size and type model removed trend in fuel effeciency",
       subtitle = "There is more variation for smaller emgines",
       x = "Displacement (L)",
       y = "Residuals after Car Class Influence Removed")

```


3.Take an exploratory graphic that you’ve created in the last month, and add informative titles to make it easier for others to understand.

```{r}

########################################
# map tows
########################################

# load shapefile of coast
shapefile <- readOGR(here::here("Rmarkdown", "Input", "Spatial"), "Land", verbose = FALSE)

# load tow data
dry_gsi <- read_csv(here::here("Rmarkdown", "Input", "dry_gsi_loc.csv"),
                    col_types =   cols(
                      .default = col_double(),
                      FISH_NUMBER = col_character()))

# convert to a dataframe for use in ggplot2
shapefile_df <- fortify(shapefile)

# get map of north america
nAmerica <- map_data("world") %>% 
  filter(region %in% c("Canada", "USA"))

# make map of north america with bounding box
can <- ggplot() +
  geom_map(data = nAmerica, map = nAmerica, aes(long, lat, map_id = region),
           color = "black", fill = "gray90") +
  coord_quickmap() +
  xlim(-150, -55) +
  ylim(30, 70) +
  # add bounding box showing location of BC map
  geom_rect(aes(xmin = -137, xmax = -122, 
                ymax = 48, ymin = 59), 
            fill = NA, color = "red", size = 1) +
  geom_text(aes(-105, 55), label = "Canada") +
  geom_text(aes(-103, 40), label = "USA") +
    labs(x = "Longitude",
         y = "Latitude",
         title = "Map of North America with Box") +
  theme(panel.grid = element_blank(),
        strip.background = element_rect(fill = "white")) 

# Display tow locations on map
  bc <- ggplot(dry_gsi, aes(LONG, LAT)) +
  geom_path(data = shapefile_df,
            aes(long, lat, group = group),
            color = "darkgrey", na.rm = TRUE) +
  geom_point(shape = 16, color = "darkred") +
  coord_quickmap() +
  theme(panel.grid = element_blank(),
        strip.background = element_rect(fill = "white")) +
  xlim(-137, -122) +
  ylim(48, 59) +
  labs(x = "Longitude",
       y = "Latitude",
       title = "BC Coast with Salmon Tows",
         caption = "Source: HSSALMON")
  
  mapstogether <- ggarrange(can, bc, labels = "AUTO")
  
  mapstogether

```

