---
title: "28 Graphics for Communication"
date: "June 2020"
output: 
      html_document: default
      
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load libraries
library(tidyverse)

# change default graph type to one I like
theme_set(theme_bw())

# load libraries for my graphic
library(ggmap) # map of canada
library(ggpubr) # panel of plots
library(rgdal) # shapefiles


```

## 28.2.1 Label Exercises

1.Create one plot on the fuel economy data with customised title, subtitle, caption, x, y, and colour labels.

```{r}
# given but change labels
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  labs(title = "Fuel economy changes by car type",
       subtitle = "(Two seater sports cars are the exception)",
       caption = "Exercise 28.2.1",
    x = "Engine displacement (L)",
    y = "Highway fuel economy (mpg)",
    colour = "Car type"
  )

```


2.The geom_smooth() is somewhat misleading because the hwy for large engines is skewed upwards due to the inclusion of lightweight sports cars with big engines. Use your modelling tools to fit and display a better model.

```{r}

ggplot(mpg, aes(displ, hwy, color = class)) +
  geom_point(aes(color = class)) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Different fuel effciency in different car classes",
       subtitle = "All negative relationships except 2 seaters",
       caption = "Exercise 28.2.2",
    x = "Engine displacement (L)",
    y = "Highway fuel economy (mpg)",
    colour = "Car type"
  )

# try different models
mod_displ <- lm(hwy ~ displ, data = mpg)
mod_class <- lm(hwy ~ class, data = mpg)
mod_displ_class <- lm(hwy ~ displ + class, data = mpg)
mod_displ_class_interact <- lm(hwy ~ displ * class, data = mpg)

# compare models and choose smallest AIC
AIC(mod_displ, mod_class, mod_displ_class, mod_displ_class_interact)

# graph residuals
mpg %>% 
  modelr::add_residuals(., mod_displ_class_interact) %>%
  ggplot(aes(displ, resid, color = class)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(title = "Car size and type model removed trend in fuel effeciency",
       subtitle = "There is more variation for smaller emgines",
       x = "Displacement (L)",
       y = "Residuals after Car Class Influence Removed")

```


3.Take an exploratory graphic that you’ve created in the last month, and add informative titles to make it easier for others to understand.

```{r}

########################################
# map tows
########################################

# load shapefile of coast
shapefile <- readOGR(here::here("Rmarkdown", "Input", "Spatial"), "Land", verbose = FALSE)

# load tow data
dry_gsi <- read_csv(here::here("Rmarkdown", "Input", "dry_gsi_loc.csv"),
                    col_types =   cols(
                      .default = col_double(),
                      FISH_NUMBER = col_character()))

# convert to a dataframe for use in ggplot2
shapefile_df <- fortify(shapefile)

# get map of north america
nAmerica <- map_data("world") %>% 
  filter(region %in% c("Canada", "USA"))

# make map of north america with bounding box
can <- ggplot() +
  geom_map(data = nAmerica, map = nAmerica, aes(long, lat, map_id = region),
           color = "black", fill = "gray90") +
  coord_quickmap() +
  xlim(-150, -55) +
  ylim(30, 70) +
  # add bounding box showing location of BC map
  geom_rect(aes(xmin = -137, xmax = -122, 
                ymax = 48, ymin = 59), 
            fill = NA, color = "red", size = 1) +
  geom_text(aes(-105, 55), label = "Canada") +
  geom_text(aes(-103, 40), label = "USA") +
    labs(x = "Longitude",
         y = "Latitude",
         title = "Map of North America with Box") +
  theme(panel.grid = element_blank(),
        strip.background = element_rect(fill = "white")) 

# Display tow locations on map
  bc <- ggplot(dry_gsi, aes(LONG, LAT)) +
  geom_path(data = shapefile_df,
            aes(long, lat, group = group),
            color = "darkgrey", na.rm = TRUE) +
  geom_point(shape = 16, color = "darkred") +
  coord_quickmap() +
  theme(panel.grid = element_blank(),
        strip.background = element_rect(fill = "white")) +
  xlim(-137, -122) +
  ylim(48, 59) +
  labs(x = "Longitude",
       y = "Latitude",
       title = "BC Coast with Salmon Tows",
         caption = "Source: HSSALMON")
  
  mapstogether <- ggarrange(can, bc, labels = "AUTO")
  
  mapstogether

```

## 28.3.1 Exercises

1.Use geom_text() with infinite positions to place text at the four corners of the plot.

```{r}

label <- tibble(
  displ = c(Inf, -Inf, -Inf, Inf), 
  hwy = c(Inf, Inf, -Inf, -Inf),
  label = c("Top Right", "Top Left", "Bottom Left", "Bottom Right"),
  vjust = c("top", "top", "bottom", "bottom"),
  hjust = c("right", "left", "left", "right")
)

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_text(aes(label = label, vjust = vjust, hjust = hjust), data = label) 
# note that vjust and hjust need to be inside aes() to work

```


2.Read the documentation for annotate(). How can you use it to add a text label to a plot without having to create a tibble?
```{r}

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  annotate("text", x = 5.5, y = 35, label = "test annotate")

```


3.How do labels with geom_text() interact with faceting? How can you add a label to a single facet? How can you put a different label in each facet? (Hint: think about the underlying data.)

Labels are the same in all facets, or they repeat within facets causing blurry text. Use a new data frame to add labels different in facets

```{r}

# example in facets that produces labels same in all
label <- tibble(
  displ = Inf,
  hwy = Inf,
  label = "Labels in facets"
)

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  facet_wrap(~class) +
  geom_text(aes(label = label),
    data = label, vjust = "top", hjust = "right")

# try again with facets different
label <- tibble(
  displ = Inf,
  hwy = Inf,
  label = c("compact text", "Midsize text", "SUV text", "2 seater text", "Minivan text", "Pickup text", "Subcompact text"),
  class = unique(mpg$class)
)

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  facet_wrap(~class) +
  geom_text(aes(label = label),
    data = label, vjust = "top", hjust = "right")

```


4.What arguments to geom_label() control the appearance of the background box?


```{r}

label <- tibble(
  displ = c(Inf, -Inf, -Inf, Inf), 
  hwy = c(Inf, Inf, -Inf, -Inf),
  label = c("Top Right", "Top Left", "Bottom Left", "Bottom Right"),
  vjust = c("top", "top", "bottom", "bottom"),
  hjust = c("right", "left", "left", "right")
)

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_label(aes(label = label, vjust = vjust, hjust = hjust), data = label, color = "red",
              label.padding = unit(0.5, "lines"), # space around labels; default 0.25 lines
              label.r = unit(0.3, "lines"), # rounded corners; default 0.15 lines
              label.size = 0.3) # size of border in mm; 0.25 derfault) 

# only works with geom_label (not geom_text)
# I would use other methods such as annotate (few labels) or geom_label_repel (many labels)

```

5.What are the four arguments to arrow()? How do they work? Create a series of plots that demonstrate the most important options.

arrow specifies arrow head on geom_segment and geom_curve

angle
The angle of the arrow head in degrees (smaller numbers produce narrower, pointier arrows). Essentially describes the width of the arrow head.

length	
A unit specifying the length of the arrow head (from tip to base).

ends	
One of "last", "first", or "both", indicating which ends of the line to draw arrow heads.

type	
One of "open" or "closed" indicating whether the arrow head should be a closed triangle.


```{r}

df = tibble(
  x1 = 3,
  x2 = 5,
  y1 = 25,
  y2 = 16
)

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  labs(subtitle = "curve with default arrow") +
  geom_curve(aes(x = x1, y = y1, xend = x2, yend = x2),
             data = df,
             arrow = arrow())

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  labs(subtitle = "curve with arrow length") +
  geom_curve(aes(x = x1, y = y1, xend = x2, yend = x2),
             data = df,
             arrow = arrow(length = unit(5, "mm")))

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  labs(subtitle = "curve with arrow length and angle") +
  geom_curve(aes(x = x1, y = y1, xend = x2, yend = x2),
             data = df,
             arrow = arrow(length = unit(5, "mm"),
                           angle = 70)
             )

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  labs(subtitle = "curve with arrow length, angle, and end") +
  geom_curve(aes(x = x1, y = y1, xend = x2, yend = x2),
             data = df,
             arrow = arrow(length = unit(5, "mm"),
                           angle = 70,
                           ends = "both")
             )

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  labs(subtitle = "curve with arrow length, angle, end, and type") +
  geom_curve(aes(x = x1, y = y1, xend = x2, yend = x2),
             data = df,
             arrow = arrow(length = unit(5, "mm"),
                           angle = 70,
                           ends = "last",
                           type = "closed")
             )



```

