---
title: "3 Data Visualization"
author: "Erika Anderson"
date: "March 15, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# this is only necessary if you haven't used tidyverse ever
# install.packages("tidyverse")

# load library
library(tidyverse)
```

### R Visualization Exercises

 See section [3 Data Visualisation](https://r4ds.had.co.nz/data-visualisation.html#exercises-4) for text and exercises. I found some answers [online](https://rpubs.com/phi/271994).
 
##3.2.4 Exercises

 # 1. Run 
```{r 3.2.4.1}
ggplot(data = mpg)

```
      
      What do you see? 
       - Empty plot

 # 2. mpg dataset
```{r 3.2.4.2}
# show data set
mpg

```
      
      Use ?mpg for more information
    
      How many rows are in mpg? 
        - 234 
      How many columns? 
      - 11
      mpg
      
 # 3. What does the drv variable describe? Read the help for ?mpg to find out.

      drive train: f = front-wheel drive, r = rear wheel drive, 4 = 4wd

 # 4. Make a scatterplot of hwy vs cyl.
 
```{r 3.2.4.4}

      ggplot(data = mpg) +
        geom_point(aes(x = hwy, y = cyl))

```
 
 # 5. What happens if you make a scatterplot of class vs drv? 
```{r 3.2.4.5}
      ggplot(data = mpg) +
        geom_point(aes(x = class, y = drv))

```
 
 Why is the plot not useful?
   - categorical variables so need different plot type

