---
title: "4 Workflow & 5 Data Transformation"
author: "Erika Anderson"
date: "2019-03-27"
output: html_document
---

A group of DFO coworkers decided to work together to compete the exercises in [R for Data Science](https://r4ds.had.co.nz/) by Garrett Grolemund and Hadley Wickham. Many of the solutions to the exercises are available online, with the most [extensive version](https://jrnold.github.io/r4ds-exercise-solutions/index.html) by Jeffrey Arnold on Git Hub.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load library
library(tidyverse)
```

# 4 Workflow Exercises

 See section [4 Workflow: basics](https://r4ds.had.co.nz/workflow-basics.html) for text and exercises. There are  [online answers](https://jrnold.github.io/r4ds-exercise-solutions/transform.html#add-new-variables-with-mutate) available for most of the exercises.
 
## 4.4 Practices Exercises

### 4.4.1 Why does this code not work?
 
 
```{r 4.4.1}

my_variable <- 10
my_varıable

```

There is a l instead of a i within the variable name.

### 4.4.2 Tweak each of the following R commands so that they run correctly:

```{r 4.4.2a}

# library(tidyverse) 
# do not need since loaded in earlier code chunk

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

#fliter(mpg, cyl = 8)

```

- correct misspelled function name filter
- use double equal symbols for filter and not assignment

```{r 4.4.2b}

filter(mpg, cyl == 8)

```

```{r 4.4.2c}

#filter(diamond, carat > 3)

```
- the data set is not called diamond

```{r 4.4.2d}

 filter(diamonds, carat > 3)

```

Press Alt + Shift + K. What happens? 

    RStudio opens a keyboard shortcut menu.

How can you get to the same place using the menus?

    Help Menu -> Keyboard Shortcuts Help

# 5 Data Transformation Exercises

 See section [5 Data Tranformation](https://r4ds.had.co.nz/transform.html) for text and exercises. Answers are available [online](https://jrnold.github.io/r4ds-exercise-solutions/workflow-basics.html).
 
```{r 5_data}

# load library with data
library(nycflights13)

flights

```

## 5.2.4 Exercises

### 1.Find all flights that

    1.Had an arrival delay of two or more hours 
```{r}
    filter(flights, arr_delay >= 120)
```
    
    2.Flew to Houston (IAH or HOU)
```{r}
filter(flights, dest == "IAH" | dest == "HOU")
filter(flights, dest %in% c("IAH", "HOU"))

```

    3.Were operated by United, American, or Delta
```{r}

unique(flights$carrier)

filter(flights, carrier %in% c("UA", "AA", "DL"))
```
    
    4.Departed in summer (July, August, and September)
```{r}
  filter(flights, month %in% c(7 , 8 , 9))

```
    
    5.Arrived more than two hours late, but didn’t leave late
```{r}
  filter(flights, arr_delay > 120 & dep_delay <= 0)
```
    
    6.Were delayed by at least an hour, but made up over 30 minutes in flight
```{r}
  filter(flights, dep_delay > 60 & dep_delay - arr_delay >= 30)
```
    
    7.Departed between midnight and 6am (inclusive)
```{r}
flights %>%
  filter(dep_time >= 000 & dep_time <= 600)

```
 
### 2. Another useful dplyr filtering helper is between(). What does it do? 

You can filter all values between or equal to two values using the between function.

### Can you use it to simplify the code needed to answer the previous challenges?

    4.Departed in summer (July, August, and September)
```{r}
  filter(flights, between(month, 7,9))

```
 
    7.Departed between midnight and 6am (inclusive)
```{r}
  filter(flights, between(dep_time, 000, 600))

```

### 3. How many flights have a missing dep_time? What other variables are missing? What might these rows represent?
```{r}
filter(flights, is.na(dep_time))

nrow(filter(flights, is.na(dep_time)))
```

These flights may have been cancelled since they have missing dep_time, dep_delay, arr_time etc.

### 4. Why is NA ^ 0 not missing? 

Any number to the power of zero is one.

### Why is NA | TRUE not missing? 

Because NA or true is not missing

### Why is FALSE & NA not missing? 

Because FALSE and NA is FALSE

### Can you figure out the general rule? (NA * 0 is a tricky counterexample!)
If you use NA in a calculation it becomes NA (except to the power of zero).

## 5.3.1 Exercises
####How could you use arrange() to sort all missing values to the start? (Hint: use is.na()).
```{r}
arrange(flights, desc(is.na(dep_time)))
```


###Sort flights to find the most delayed flights. Find the flights that left earliest.
```{r}
arrange(flights, desc(dep_delay))

arrange(flights, dep_delay)
```


### Sort flights to find the fastest flights.
```{r}
arrange(flights, air_time)

```


### Which flights travelled the longest? Which travelled the shortest?
```{r}
arrange(flights, desc(air_time))

arrange(flights, air_time)
```

The longest flights were 11.5 hours and the shortest flights were 20 minutes.

## 5.4.1 Exercises

### Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.
```{r}
select(flights, dep_time, dep_delay, arr_time, arr_delay)

select(flights, dep_time, arr_time, ends_with("delay"))

select(flights, dep_time, arr_time, contains("delay"))
```


### What happens if you include the name of a variable multiple times in a select() call?
```{r}
select(flights, year, year, month, month,)

```

Duplicate column names does not appear to change the select results at all.


### What does the one_of() function do? Why might it be helpful in conjunction with this vector?
```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")

select(flights, one_of(vars))

```

one_of is a helper function that allows a vector of column names as input into the select function.


### Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?

```{r}
select(flights, contains("TIME"))
# ignore.case = TRUE is the default so all column with time are selected regardless of case

# no column are selected if you change the default
select(flights, contains("TIME", ignore.case = FALSE))


```
##5.5.2 Exercises
### 1.Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.

```{r}

# select smaller data frame
flights_sml <- select(flights, time_hour, dep_time, sched_dep_time)

# create new columns
  mutate(flights_sml, dep_time_cont = (dep_time %/% 100) * 60 +  (dep_time %% 100) %% 1440,
                      sched_dep_time_cont = (sched_dep_time %/% 100) * 60 +  (sched_dep_time %% 100) %% 1440)
  

```

### 2.Compare air_time with arr_time - dep_time. What do you expect to see? What do you see? What do you need to do to fix it?

```{r}
flights %>%
  select(., air_time, arr_time, dep_time) %>%
  mutate(., arr_time_cont = (arr_time %/% 100) * 60 +  (arr_time %% 100) %% 1440,
            dep_time_cont = (dep_time %/% 100) * 60 +  (dep_time %% 100) %% 1440,
            diff = arr_time_cont - dep_time_cont,
            diff2 = air_time - diff) %>%
  arrange(desc(diff2)) %>%
  filter(., !is.na(diff2)) %>%

ggplot() +
  geom_point(aes(air_time, diff))
```

I expected to see that air time equals the difference between arrival and departure times, BUT they are not equal. In fact, there are significant differences with a pattern. The help file on the flights data set says that arr_time and dep_time are in local time zone values. It should be fixed by converting to universal time.

### 3. Compare dep_time, sched_dep_time, and dep_delay. How would you expect those three numbers to be related?

I would expect that scheduled departure time plus departure delay should equal then departure time.
```{r}

flights %>%
  select(., dep_time, sched_dep_time, dep_delay) %>%
  mutate(., sched_dep_time_cont = (sched_dep_time %/% 100) * 60 +  (sched_dep_time %% 100) %% 1440,
            dep_time_cont = (dep_time %/% 100) * 60 +  (dep_time %% 100) %% 1440,
         dep_time_calc = sched_dep_time_cont + dep_delay,
          diff = dep_time_calc - dep_time_cont) %>%
  filter(., diff != 0) %>%
  arrange(desc(diff))
```
All these results have 24 hours difference or 1440 minutes so the relationship does hold, although you have to correct for flights delayed over midnight.

### 4. Find the 10 most delayed flights using a ranking function. How do you want to handle ties? Carefully read the documentation for min_rank().

```{r}
flights %>%
  mutate(., rank = min_rank(dep_delay)) %>%
  arrange(desc(rank)) %>%
  slice(1:10)
```
There are no ties in this results, but the default for min_rank() is ties.method = "min". This means the min rank to the tied values is assigned to all tied values and the ranking restarts at the next value after accounting for ties (ie. ranks 1,2,2,2,5).

### 5.What does 1:3 + 1:10 return? Why?
```{r}
 1:3 + 1:10
```
This is warning that the vectors are unequal length and are being recycled. This can give incorrect results if that is not your intention.

### 6. What trigonometric functions does R provide?

Look at ?Trig to see all trigonometric functions in R or Google the ones you are interested in...


