---
title: "5 Data Transformation & 6 Workflow: scripts"
author: "Erika Anderson"
date: "2019-04-10"
output: html_document
---

A group of DFO coworkers decided to work together to compete the exercises in [R for Data Science](https://r4ds.had.co.nz/) by Garrett Grolemund and Hadley Wickham. Many of the solutions to the exercises are available online, with the most [extensive version](https://jrnold.github.io/r4ds-exercise-solutions/index.html) by Jeffrey Arnold on Git Hub.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load library
library(tidyverse)
```

# 5 Data Transformation Exercises continued

 See section [5 Data Tranformation](https://r4ds.had.co.nz/transform.html) for text and exercises. Answers are available [online](https://www.patrickjomalley.com/2017/10/data-transformation/#grouped-summaries-with-summarise).
 
```{r 5_data}

# load library with data
library(nycflights13)

flights

```

## 5.6.1 Exercises

### 5.6.1.1.Brainstorm at least 5 different ways to assess the typical delay characteristics of a group of flights. Consider the following scenarios:

* A flight is 15 minutes early 50% of the time, and 15 minutes late 50% of the time.
```{r}
    flights %>%
  group_by(flight) %>%
  summarise(early_15_min = sum(arr_delay <= -15, na.rm = TRUE) / n(),
            late_15_min = sum(arr_delay >= 15, na.rm = TRUE) / n()) %>%
  filter(early_15_min == 0.5 & late_15_min == 0.5)

```

* A flight is always 10 minutes late.
```{r}
flights %>%
  group_by(flight) %>%
  summarise(propLate = n_distinct(arr_delay, na.rm = TRUE)/ n(),
            mean_arr_delay = mean(arr_delay, na.rm = TRUE),
            n = n()) %>%
  filter(propLate ==1 & mean_arr_delay == 10)
    
```

* A flight is 30 minutes early 50% of the time, and 30 minutes late 50% of the time.
```{r}
flights %>%
  group_by(flight) %>%
  summarise(early30 = sum(arr_delay <= -30, na.rm = TRUE)/n(),
            late30 = sum(arr_delay <= 30, na.rm = TRUE)/n()) %>%
filter(early30 == 0.5 & late30 == 0.5)
    
```

* 99% of the time a flight is on time. 1% of the time it’s 2 hours late.
```{r}
flights %>%
  group_by(flight) %>%
  summarise(ontime = sum(arr_delay == 0, na.rm = TRUE)/n(),
            verylate = sum(arr_delay <= 120, na.rm = TRUE)/n()) %>%
filter(ontime == 0.9 & verylate == 0.01)
  
    
```

Which is more important: arrival delay or departure delay?

Arrival delay is more important since it affects making connections and schedules. Departure delay only affect wait time in the ariport


### 5.6.7.2.Come up with another approach that will give you the same output as not_cancelled %>% count(dest) and not_cancelled %>% count(tailnum, wt = distance) (without using count()).

```{r}

# start with same dataset
not_cancelled <- flights %>% 
        filter(!is.na(dep_time))

# same as not_cancelled %>% count(dest)
not_cancelled %>%
  group_by(dest) %>%
  summarise(sumDest = n())

# same as count(tailnum, wt = distance)
not_cancelled %>%
  group_by(tailnum) %>%
  summarise(ndistance = sum(distance))

```

### 5.6.7.3 Our definition of cancelled flights (is.na(dep_delay) | is.na(arr_delay) ) is slightly suboptimal. Why? Which is the most important column?

```{r}
flights %>% 
  filter(is.na(dep_delay) & !is.na(arr_delay))
```

There are no flights that have NA arrival and departure since if a flight departed it has to land (unless their is a rare accident). Departure column is most important then.


### 5.6.7.4. Look at the number of cancelled flights per day. Is there a pattern? Is the proportion of cancelled flights related to the average delay?

```{r}
flights %>%
  group_by(year, month, day) %>%
  summarise(cancelled = sum(is.na(dep_time)),
            avg_delay = mean(arr_delay, na.rm = TRUE)) %>%
  
  ggplot(aes(log(cancelled), avg_delay)) +
    geom_point() +
  geom_smooth(method = "lm")
```

It looks like more flights are cancelled when there are longer delays. Perhaps due to weather influencing airports.

#### 5.6.7.5. Which carrier has the worst delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about flights %>% group_by(carrier, dest) %>% summarise(n()))

```{r}
flights %>%
  group_by(carrier) %>%
  summarise(avg_delay = mean(arr_delay, na.rm = TRUE)) %>%
  arrange(desc(avg_delay))

flights %>%
  group_by(dest) %>%
  summarise(avg_delay = mean(arr_delay, na.rm = TRUE)) %>%
  arrange(desc(avg_delay))

flights %>%
  group_by(carrier, dest) %>%
  summarise(avg_delay = mean(arr_delay, na.rm = TRUE)) %>%
  arrange(desc(avg_delay))
  
```

It is hard to seperate out both carrier and destination effects using summarise. There was an online suggestion to filter out all the destinations with long delays, then rank the carriers.

### 5.6.7.6. What does the sort argument to count() do. When might you use it?

  It sorts results into descending order.

```{r}
flights %>% 
        count(dest, sort = TRUE)
```


### 5.7.1.1 Refer back to the lists of useful mutate and filtering functions. Describe how each operation changes when you combine it with grouping.

  When grouped they all apply the function to the grouped variable.

```{r}

x <- c(10:20)

mean(x)
median(x)
sd(x) # standard deviation
IQR(x) # interquartile range
mad(x) # median absolute deviation
min(x)
max(x)
quantile(x, 0.25)
first(x)
nth(x, 2)
last(x)
#count(x)
#n(x)
n_distinct(x)
sum(x)


```


### 5.7.1.2 Which plane (tailnum) has the worst on-time record?

```{r}
flights %>%
  group_by(tailnum) %>%
   summarise(., max_delay = max(arr_delay, na.rm = TRUE)) %>% 
  arrange(desc(max_delay)) %>%
   top_n(1, max_delay)
  
```


### 5.7.1.3. What time of day should you fly if you want to avoid delays as much as possible?

  Traqvel in the morning to reduce delays.

```{r}
flights %>% 
        group_by(hour) %>% 
        summarise(avg.delay = mean(arr_delay, na.rm = TRUE)) %>% 
        arrange(avg.delay)

```


### 5.7.1.4. For each destination, compute the total minutes of delay. For each flight, compute the proportion of the total delay for its destination.

```{r}

flights %>% 
        group_by(dest) %>% 
        filter(arr_delay > 0) %>% 
        mutate(tot.delay = sum(arr_delay),
               prop.delay = arr_delay / tot.delay) %>% 
        select(year:day, dest, arr_delay, tot.delay, prop.delay)
  

```


### 5.7.1.5 Delays are typically temporally correlated: even once the problem that caused the initial delay has been resolved, later flights are delayed to allow earlier flights to leave. Using lag(), explore how the delay of a flight is related to the delay of the immediately preceding flight.

```{r}
flights %>% 
        group_by(origin) %>% 
        arrange(year, month, day, dep_time) %>% 
        filter(!is.na(dep_delay)) %>% 
        mutate(prev.delay = lag(dep_delay)) %>%
  
  # graph it
  ggplot(aes(dep_delay, prev.delay)) +
  geom_point()
```



### 5.7.1.6. Look at each destination. Can you find flights that are suspiciously fast? (i.e. flights that represent a potential data entry error). Compute the air time a flight relative to the shortest flight to that destination. Which flights were most delayed in the air?

```{r}
flights %>% 
        group_by(dest) %>% 
        mutate(avg.airtime = mean(air_time, na.rm = TRUE), scl_airtime = scale(air_time), n = n()) %>% 
        select(year:day, dest, air_time, avg.airtime, scl_airtime, n) %>% 
        filter(scl_airtime < -4) %>%
        arrange(desc(scl_airtime))
```


### 5.7.1.7. Find all destinations that are flown by at least two carriers. Use that information to rank the carriers.

```{r}
flights_2dest <- flights %>% 
        group_by(dest) %>% 
        summarise(num.carrier = length(unique(carrier))) %>% 
        filter(num.carrier >= 2) %>% 
        arrange(desc(num.carrier))

flights %>%
  filter(dest %in% flights_2dest$dest) %>%
  group_by(carrier) %>%
  summarise(numDest = sum(length(unique(dest)))) %>%
  arrange(desc(numDest))


```


### 5.7.1.8. For each plane, count the number of flights before the first delay of greater than 1 hour.

```{r}
flights %>%
  group_by(tailnum) %>%
  summarise(firstLate = first(arr_delay > 60))

flights %>%
  arrange(tailnum, year, month, day) %>%
  group_by(tailnum) %>%
  mutate(delay_1hr = dep_delay > 60) %>%
  mutate(before_delay = cumsum(delay_1hr)) %>%
  filter(before_delay < 1) %>%
  count(sort = TRUE)
```

# Workflow: scripts

## 6.3 Practice

### 6.3.1 Go to the RStudio Tips twitter account, https://twitter.com/rstudiotips and find one tip that looks interesting. Practice using it!

Ctrl + Alt + B run code from the beginning  
Crtl + I fix indentation  

### 6.3.2 What other common mistakes will RStudio diagnostics report? Read https://support.rstudio.com/hc/en-us/articles/205753617-Code-Diagnostics to find out.

whitespace  
unused variable  
using a variable before defining it  
expected commas  
undefined function  

