---
title: "5 Data Transformation & 6 Workflow: scripts"
author: "Erika Anderson"
date: "2019-04-10"
output: html_document
---

A group of DFO coworkers decided to work together to complete the exercises in [R for Data Science](https://r4ds.had.co.nz/) by Garrett Grolemund and Hadley Wickham. Many of the solutions to the exercises are available online, with the most [extensive version](https://jrnold.github.io/r4ds-exercise-solutions/index.html) by Jeffrey Arnold on Git Hub.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load library
library(tidyverse)
library(magrittr)
```

# 5 Data Transformation Exercises continued

 See section [5 Data Tranformation](https://r4ds.had.co.nz/transform.html) for text and exercises. Answers are available [online](https://www.patrickjomalley.com/2017/10/data-transformation/#grouped-summaries-with-summarise).
 
```{r 5_data}

# load library with data
library(nycflights13)

flights

```

## 5.6.1 Exercises

### 5.6.1.1.Brainstorm at least 5 different ways to assess the typical delay characteristics of a group of flights. Consider the following scenarios:

* A flight is 15 minutes early 50% of the time, and 15 minutes late 50% of the time.
```{r}

  # Incorrectly include NAs in denominator so Micheal's way is better
    flights %>%
  group_by(flight) %>%
  summarise(early_15_min = sum(arr_delay <= -15, na.rm = TRUE) / n(),
            late_15_min = sum(arr_delay >= 15, na.rm = TRUE) / n()) %>%
  filter(early_15_min == 0.5 & late_15_min == 0.5)

# Micheal's answer
    flights %>%
  group_by(flight) %>%
  summarise(early_15_min = mean(arr_delay <= -15, na.rm = TRUE),
            late_15_min = mean(arr_delay >= 15, na.rm = TRUE)) %>%
  filter(early_15_min == 0.5 & late_15_min == 0.5)
```

* A flight is always 10 minutes late.
```{r}
flights %>%
  group_by(flight) %>%
  summarise(propLate = n_distinct(arr_delay, na.rm = TRUE)/n(),
            mean_arr_delay = mean(arr_delay, na.rm = TRUE),
            n = n()) %>%
  filter(propLate == 1 & mean_arr_delay == 10)

# Michael's way
flights %>%
  group_by(flight) %>%
  filter(!is.na(arr_delay)) %>%
  summarise(propLate = mean(arr_delay == 10, na.rm = TRUE)) %>%
  filter(propLate == 1)
    
```

* A flight is 30 minutes early 50% of the time, and 30 minutes late 50% of the time.
```{r}
flights %>%
  group_by(flight) %>%
  summarise(early30 = sum(arr_delay <= -30, na.rm = TRUE)/n(),
            late30 = sum(arr_delay <= 30, na.rm = TRUE)/n()) %>%
filter(early30 == 0.5 & late30 == 0.5)
    
```

* 99% of the time a flight is on time. 1% of the time it’s 2 hours late.
```{r}
flights %>%
  group_by(flight) %>%
  summarise(ontime = sum(arr_delay == 0, na.rm = TRUE)/n(),
            verylate = sum(arr_delay <= 120, na.rm = TRUE)/n()) %>%
filter(ontime == 0.9 & verylate == 0.01)

  
    
```

Nick remindered us about & and &&. Use & if using row by row, but && needed if there is a single results that needs to be recycled. Such as if_else with one condition.


Which is more important: arrival delay or departure delay?

Arrival delay is more important since it affects making connections and schedules. Departure delay only affect wait time in the ariport


### 5.6.7.2.Come up with another approach that will give you the same output as not_cancelled %>% count(dest) and not_cancelled %>% count(tailnum, wt = distance) (without using count()).

```{r}

# start with same dataset
not_cancelled <- flights %>% 
        filter(!is.na(dep_time))

# same as not_cancelled %>% count(dest)
not_cancelled %>%
  group_by(dest) %>%
  summarise(sumDest = n())

# same as count(tailnum, wt = distance)
not_cancelled %>%
  group_by(tailnum) %>%
  summarise(ndistance = sum(distance))

```

### 5.6.7.3 Our definition of cancelled flights (is.na(dep_delay) | is.na(arr_delay) ) is slightly suboptimal. Why? Which is the most important column?

```{r}

# no results
flights %>% 
  filter(is.na(dep_delay) & !is.na(arr_delay))

# opposite view with more results
flights %>% 
  filter(!is.na(dep_delay) & is.na(arr_delay))
```

There are no flights that have NA arrival and departure since if a flight departed it has to land (unless their is a rare accident). Departure column is most important then.


### 5.6.7.4. Look at the number of cancelled flights per day. Is there a pattern? Is the proportion of cancelled flights related to the average delay?

```{r}
flights %>%
  group_by(year, month, day) %>%
  summarise(cancelled = sum(is.na(dep_time)),
            avg_delay = mean(arr_delay, na.rm = TRUE)) %>%
  
  ggplot(aes(cancelled, avg_delay)) +
    geom_point() +
  geom_smooth() 

# Nick's aside
# use vector function on data frame columns
flights %$%
  sum(arr_delay, dep_delay, na.rm = TRUE)

```

It looks like more flights are cancelled when there are longer delays. Perhaps due to weather influencing airports.

#### 5.6.7.5. Which carrier has the worst delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about flights %>% group_by(carrier, dest) %>% summarise(n()))

```{r}

#
flights %>%
  group_by(carrier) %>%
  summarise(avg_delay = mean(arr_delay, na.rm = TRUE)) %>%
  arrange(desc(avg_delay))

flights %>%
  group_by(dest) %>%
  summarise(avg_delay = mean(arr_delay, na.rm = TRUE)) %>%
  arrange(desc(avg_delay))

flights %>%
  group_by(carrier, dest) %>%
  summarise(avg_delay = mean(arr_delay, na.rm = TRUE)) %>%
  arrange(desc(avg_delay))

```

It is hard to seperate out both carrier and destination effects using summarise. There was an online suggestion to filter out all the destinations with long delays, then rank the carriers.

Jocelyn used an [answer](http://nickg.bio/_posts/ngiangre/r_for_data_science_solutions.html#5_which_carrier_has_the_worst_delays_challenge:_can_you_disentangle_the_effects_of_bad_airports_vs_bad_carriers_whywhy_not_(hint:_think_about_flights_%%_group_by(carrier,_dest)_%%_summarise(n()))) from Nick G.

```{r}

cancelled_delayed <- 
  flights %>%
  mutate(cancelled = (is.na(arr_delay) | is.na(dep_delay))) %>%
  group_by(carrier) %>%
  summarise(avg_dep_delay = mean(dep_delay, na.rm = TRUE),
            avg_arr_delay = mean(arr_delay, na.rm = TRUE))  

#departure delay barplot
ggplot(cancelled_delayed, aes(x = carrier, y = avg_dep_delay)) +
  geom_bar(stat="identity") 

#arrival delay barplot
ggplot(cancelled_delayed, aes(x = carrier, y = avg_arr_delay)) +
  geom_bar(stat="identity") 


airports_delayed <- 
  flights %>%
  mutate(cancelled = (is.na(arr_delay) | is.na(dep_delay))) %>%
  group_by(carrier, origin) %>%
  summarise(avg_dep_delay = mean(dep_delay, na.rm = TRUE),
            avg_arr_delay = mean(arr_delay, na.rm = TRUE))  

#facet wrapped by carrier, coloured by airport of departure
ggplot(airports_delayed, aes(x = factor(1), y = avg_dep_delay, fill = origin)) +
  geom_bar(stat="identity")+
  facet_wrap(~carrier)




#looking at destination airport instead
airports_delayed_dest <- 
  flights %>%
  mutate(cancelled = (is.na(arr_delay) | is.na(dep_delay))) %>%
  group_by(carrier, dest) %>%
  summarise(avg_dep_delay = mean(dep_delay, na.rm = TRUE),
            avg_arr_delay = mean(arr_delay, na.rm = TRUE))  

#facet wrapped by carrier, coloured by airport of departure
ggplot(airports_delayed_dest, aes(x = factor(1), y = avg_arr_delay, fill = dest)) +
  geom_bar(stat="identity")+
  facet_wrap(~carrier)+
  theme(legend.position="none")

```


### 5.6.7.6. What does the sort argument to count() do. When might you use it?

  It sorts results into descending order.

```{r}

# using option to sort descending
flights %>% 
        count(dest, sort = TRUE)

# longer way that works too
flights %>% 
        count(dest) %>%
        arrange(desc(n))

```


### 5.7.1.1 Refer back to the lists of useful mutate and filtering functions. Describe how each operation changes when you combine it with grouping.

  When grouped they all apply the function to the grouped variable.

```{r}

x <- c(10:20)

mean(x)
median(x)
sd(x) # standard deviation
IQR(x) # interquartile range
mad(x) # median absolute deviation
min(x)
max(x)
quantile(x, 0.25)
first(x)
nth(x, 2)
last(x)
#count(x)
#n(x)
n_distinct(x)
sum(x)


```


### 5.7.1.2 Which plane (tailnum) has the worst on-time record?

```{r}

# method using summarise
flights %>%
  group_by(tailnum) %>%
   summarise(., max_delay = max(arr_delay, na.rm = TRUE)) %>% 
  arrange(desc(max_delay)) %>%
   top_n(1, max_delay)

# simplier code without group_by
flights %>%
  filter(rank(desc(arr_delay)) == 1) %>%
  select(tailnum, arr_delay)

# or you could take longest average delay as different interpretation
  
```


### 5.7.1.3. What time of day should you fly if you want to avoid delays as much as possible?

  Travel in the morning to reduce delays.

```{r}
flights %>% 
        group_by(hour) %>% 
        summarise(avg.delay = mean(arr_delay, na.rm = TRUE)) %>% 
        arrange(avg.delay)

# Jocelyn's found this without summarise
# Hmmmm not sure about this one
flights %>%
  filter(arr_delay > 0) %>%
  group_by(dest) %>%
  mutate(., arr_delay_total = sum(arr_delay),
            arr_delay_prop = arr_delay/arr_delay_total) %>%
  arrange(desc(arr_delay_total))
```


### 5.7.1.4. For each destination, compute the total minutes of delay. For each flight, compute the proportion of the total delay for its destination.

```{r}

flights %>% 
        group_by(dest) %>% 
        filter(arr_delay > 0) %>% 
        mutate(tot.delay = sum(arr_delay),
               prop.delay = arr_delay / tot.delay) %>% 
        select(year:day, dest, arr_delay, tot.delay, prop.delay)

```


### 5.7.1.5 Delays are typically temporally correlated: even once the problem that caused the initial delay has been resolved, later flights are delayed to allow earlier flights to leave. Using lag(), explore how the delay of a flight is related to the delay of the immediately preceding flight.

```{r}
flights %>% 
        group_by(origin) %>% 
        arrange(year, month, day, dep_time) %>% 
        filter(!is.na(dep_delay)) %>% 
        mutate(prev.delay = lag(dep_delay)) %>%
  
  # graph it
  ggplot(aes(dep_delay, prev.delay)) +
  geom_point() +
  geom_smooth()

# Jocelyn's version of same thing
# is dataset already arranged by year, month, day?
flights %>%
  group_by(year, month, day) %>%
  filter(!is.na(dep_delay)) %>%
  mutate(lag_delay = lag(dep_delay)) %>%
  filter(!is.na(lag_delay)) %>%
  
    # graph it
  ggplot(aes(dep_delay, lag_delay)) +
  geom_point() +
  geom_smooth()

```



### 5.7.1.6. Look at each destination. Can you find flights that are suspiciously fast? (i.e. flights that represent a potential data entry error). Compute the air time a flight relative to the shortest flight to that destination. Which flights were most delayed in the air?

```{r}
flights %>% 
        group_by(dest) %>% 
        mutate(avg.airtime = mean(air_time, na.rm = TRUE), 
               scl_airtime = scale(air_time), 
               n = n()) %>% 
        select(year:day, dest, air_time, avg.airtime, scl_airtime, n) %>% 
        filter(scl_airtime < -4) %>%
        arrange(desc(scl_airtime))

# another suggestion
# hmmmm
flights %>%
  group_by(dest) %>%
  mutate(prop_air_time = min(arr_time)/mean(arr_time)) %>%
  filter(rank(prop_air_time) >= 5) %>%
  arrange(dest, prop_air_time)
```


### 5.7.1.7. Find all destinations that are flown by at least two carriers. Use that information to rank the carriers.

```{r}
flights_2dest <- flights %>% 
        group_by(dest) %>% 
        summarise(num.carrier = length(unique(carrier))) %>% 
        filter(num.carrier >= 2) %>% 
        arrange(desc(num.carrier))

flights %>%
  filter(dest %in% flights_2dest$dest) %>%
  group_by(carrier) %>%
  summarise(numDest = sum(length(unique(dest)))) %>%
  arrange(desc(numDest))

# using n_distinct
flights %>%
  filter(dest %in% flights_2dest$dest) %>%
  group_by(carrier) %>%
  summarise(numDest = sum(n_distinct(dest))) %>%
  arrange(desc(numDest))


```


### 5.7.1.8. For each plane, count the number of flights before the first delay of greater than 1 hour.

```{r}

# online answer; not sure about cumsum()
flights %>%
  arrange(tailnum, year, month, day) %>%
  group_by(tailnum) %>%
  mutate(delay_1hr = dep_delay > 60) %>%
  mutate(before_delay = cumsum(delay_1hr)) %>%
  filter(before_delay < 1) %>%
  count(sort = TRUE)

# Nick suggested making a date column and using lubridate
# or using time_hour using two data frames

  
```

# Workflow: scripts

## 6.3 Practice

### 6.3.1 Go to the RStudio Tips twitter account, https://twitter.com/rstudiotips and find one tip that looks interesting. Practice using it!

Ctrl + Alt + B run code from the beginning  
Ctrl + A selects file
Crtl + I fix indentation
Alt +Shift + Arrow Down selects mulitple rows (flips screen upside down)
Alt + mouse click to select multiple lines


### 6.3.2 What other common mistakes will RStudio diagnostics report? Read https://support.rstudio.com/hc/en-us/articles/205753617-Code-Diagnostics to find out.

whitespace  
unused variable  
using a variable before defining it  
expected commas  
undefined function  

