---
title: "7 Exploratory Data Analysis"
author: "Erika Anderson"
date: "April 24, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load library
library(tidyverse)

```

A group of DFO coworkers decided to work together to complete the exercises in [R for Data Science](https://r4ds.had.co.nz/) by Garrett Grolemund and Hadley Wickham. Many of the solutions to the exercises are available online, with the most [extensive version](https://jrnold.github.io/r4ds-exercise-solutions/index.html) that I have found so far by Jeffrey Arnold on Git Hub.


# 7.3.4 Exercises

1. Explore the distribution of each of the x, y, and z variables in diamonds. What do you learn? Think about a diamond and how you might decide which dimension is the length, width, and depth.

```{r}

ggplot(diamonds, mapping = aes(x = x)) +
  geom_histogram(binwidth = 0.5)
```
  
Diamonds have an x dimension between approximately 4 to 8 mm with peaks at 5 and 7. The documentation says x  is "length in mm (0–10.74)"."


```{r}
ggplot(diamonds) + 
  geom_histogram(mapping = aes(x = y), binwidth = 0.5) +
  coord_cartesian(ylim = c(0, 50))
```

Diamonds have a y dimension between approximately 5 to 10 mm with outliers at 0, 30 and 60 mm. The documentation says y  is "width in mm (0–58.9)".


```{r}
ggplot(diamonds) + 
  geom_histogram(mapping = aes(x = z), binwidth = 0.5) +
  coord_cartesian(ylim = c(0, 30))
```


Diamonds have a z dimension between approximately 2 to 7 mm with outliers at 0 and 32 mm. The documentation says z is "depth in mm (0–31.8)".

2. Explore the distribution of price. Do you discover anything unusual or surprising? (Hint: Carefully think about the binwidth and make sure you try a wide range of values.)

```{r}

# default bin witdth
ggplot(diamonds, aes(x = price)) +
  geom_histogram() +
  ggtitle("More diamonds at lower prices using default bin width")

diamonds %>% 
  count(cut_width(price, 100)) %>%
  arrange(desc(n))

# turn off scientific notation
#options("scipen" = FALSE)

# change bin width
ggplot(diamonds, aes(x = price)) +
  geom_histogram(binwidth = 100) +
  ggtitle("Marketing drives prices at certain price points",
          subtitle = "Examples at $700 and $600")

```


3. How many diamonds are 0.99 carat? How many are 1 carat? What do you think is the cause of the difference?

```{r}

diammonds99 <- diamonds %>%
  filter(carat == 0.99 | carat == 1) %>%
  count(carat)
  
```

There are 23 - 0.99 carat diamonds and 1558- 1 carat diamonds. This is likely driven by marketing.

4. Compare and contrast coord_cartesian() vs xlim() or ylim() when zooming in on a histogram. What happens if you leave binwidth unset? What happens if you try and zoom so only half a bar shows?

```{r}

# basic histogram
ggplot(diamonds, aes(x = x)) +
  geom_histogram(binwidth = 0.25) 

# coord_cartesian wit xlim
ggplot(diamonds, aes(x = x)) +
  geom_histogram(binwidth = 1) +
  coord_cartesian(xlim = c(0, 6))

# xlim by itself
ggplot(diamonds, aes(x = x)) +
  geom_histogram(binwidth = 1) +
  xlim(0, 5.5)

```


Using the xlim function, there are warnings given about omitting observations. There are no warnings with coord_cartesian function. You need to choose a binwidth to omit the bin width warning for both functions. The coord_cartesian zooms into a section of the graph, whereas the xlim function drops all observations outside the limits. If the bin is cut, in coord_cartesian the in is the same height but half the width. If the bin is cut in xlim, the data is dropped in the cut bin.


# 7.4.1 Exercises

1. What happens to missing values in a histogram? What happens to missing values in a bar chart? Why is there a difference?

```{r}

diamonds2 <- diamonds %>% 
  mutate(y = ifelse(y < 3 | y > 20, NA, y))

ggplot(diamonds2, aes(x = y)) +
  geom_histogram() +
  ggtitle("Geom_histogram")

ggplot(diamonds2, aes(x = y)) +
  geom_bar() +
  ggtitle("Geom_bar")
  

```

There doesn't appear to be a large difference between geom_histogram and geom_bar because both functions give a warning of missing values. This is different with categorical data when NAs become a separate category.

geom_histogram makes bar continuous even if they are not continuous but geom_bar leaves gaps???

2. What does na.rm = TRUE do in mean() and sum()?

In the mean function, na.rm = TRUE is a logical value indicating whether NA values should be stripped before the computation proceeds.

In the sum function, na.rm = TRUE is logical. Should missing values (including NaN) be removed?

# 7.5.1.1 Exercises

1. Use what you’ve learned to improve the visualisation of the departure times of cancelled vs. non-cancelled flights.

```{r}
# load dataset
library(nycflights13)

flights %>% 
  mutate(
    cancelled = is.na(dep_time),
    sched_hour = sched_dep_time %/% 100,
    sched_min = sched_dep_time %% 100,
    sched_dep_time = sched_hour + sched_min / 60) %>%
  
  ggplot(mapping = aes(x = cancelled,  y =sched_dep_time)) + 
    geom_boxplot()

```


More flights are cancelled later in the day.


2. What variable in the diamonds dataset is most important for predicting the price of a diamond? How is that variable correlated with cut? Why does the combination of those two relationships lead to lower quality diamonds being more expensive?

```{r}

# is carat the best predictor of price>
ggplot(diamonds, aes(carat, price)) +
  geom_point() +
  ggtitle("Is carat the best predictor?")

# is clarity the best predictor of price?
ggplot(diamonds, aes(clarity, price)) +
  geom_boxplot() +
    ggtitle("Is clarity the best predictor?")
```
  
Carat is more important than clarity to predict price.

``` {r}
# how is carat related to cut?
ggplot(diamonds, aes(cut, carat)) +
  geom_boxplot() +
    ggtitle("How are carat and cut related?") +
  coord_flip()

```

Higher quality diamonds have a lesser cut than lower quality diamonds. This leads to a lower price for ideal diamonds compared to fair quality diamonds. Not much difference in the middle cut types though. Could do more on clarity too.

3. Install the ggstance package, and create a horizontal boxplot. How does this compare to using coord_flip()?

```{r}

library(ggstance)

# geom_boxplot with coord-flip
ggplot(diamonds, aes(cut, carat)) +
  geom_boxplot() +
    ggtitle("geom_boxplot") +
  coord_flip()

# ggstance
ggplot(diamonds, aes(carat, cut)) +
  geom_boxploth() +
  ggtitle("geom_boxploth")



```

Both boxplots looks the same using geom_boxplot with coord_flip or geom_boxploth.

4. One problem with boxplots is that they were developed in an era of much smaller datasets and tend to display a prohibitively large number of “outlying values”. One approach to remedy this problem is the letter value plot. Install the lvplot package, and try using geom_lv() to display the distribution of price vs cut. What do you learn? How do you interpret the plots?

```{r}

library(lvplot)

# geom_boxplot with coord-flip
ggplot(diamonds, aes(cut, carat)) +
  geom_lv(outlier.colour = "blue") +
    ggtitle("geom_lv") +
  coord_flip()

```

geom_lv uses more boxes than the quartiles used in geom_boxplot. lvplot is useful for larger datasets (>200 observations).

5. Compare and contrast geom_violin() with a facetted geom_histogram(), or a coloured geom_freqpoly(). What are the pros and cons of each method?

```{r}

ggplot(diamonds, aes(x = price, y = ..density.., color = cut)) +
  geom_freqpoly(binwidth = 100) +
  ggtitle("geom_freqpoly")

ggplot(diamonds, aes(x = price)) +
  facet_wrap(~cut) +
  geom_histogram(binwidth = 100) +
  ggtitle("facetted geom_histogram")

ggplot(diamonds, aes(x = cut, y = price)) +
  geom_violin() +
  ggtitle("geom_violin")


```

I think that geom_violin is easier to compare the three distributions. Facetted geom_histogram has better visualization of individual distributions, whereas it is easier to see each cut for a given price in geom_freqpoly.


6. If you have a small dataset, it’s sometimes useful to use geom_jitter() to see the relationship between a continuous and categorical variable. The ggbeeswarm package provides a number of methods similar to geom_jitter(). List them and briefly describe what each one does.

```{r}
library(ggbeeswarm)

## from beeswarm github website

#compare to jitter
ggplot(iris,aes(Species, Sepal.Length)) + 
  geom_jitter() +
  ggtitle("geom_jitter")

ggplot(iris,aes(Species, Sepal.Length)) +
  geom_quasirandom() +
  ggtitle("geom_quasirandom")

ggplot(iris,aes(Species, Sepal.Length)) +
  geom_beeswarm() +
  ggtitle("geom_beeswarm")


```


Looking at the example graphs it looks like geom_jitter is more random, geom_quasirandom is more like a violin plot, and geom_beeswarm is a tighter violin plot. All functions are used to offset points within categories to reduce overplotting. 
  
geom_quasirandom uses a van der Corput sequence or Tukey texturing (Tukey and Tukey "Strips displaying empirical distributions: I. textured dot strips") to space the dots to avoid overplotting.

geom_beeswarm uses the beeswarm library to do point-size based offset. 


# 7.5.2.1 Exercises

1. How could you rescale the count dataset above to more clearly show the distribution of cut within colour, or colour within cut?

```{r}

ggplot(data = diamonds) +
  geom_count(mapping = aes(x = cut, y = color)) +
  ggtitle("Original geom_count")

ggplot(data = diamonds) +
  geom_count(mapping = aes(x = cut, y = color)) +
  scale_size_area() +
  ggtitle("Rescaled geom_count")

# using geom_tile
  diamonds %>% 
  count(color, cut) %>%  
  ggplot(mapping = aes(x = color, y = cut)) +
    geom_tile(mapping = aes(fill = n)) +
    ggtitle("Original geom_tile") 
  
  # rescale using viridis   
  diamonds %>% 
  count(color, cut) %>%  
  ggplot(mapping = aes(x = color, y = cut)) +
    geom_tile(mapping = aes(fill = n)) +
    ggtitle("Viridis geom_tile") +
    scale_fill_viridis_c()
  
  
  

```

Another way would be to transform the data to emphasize differences.


2. Use geom_tile() together with dplyr to explore how average flight delays vary by destination and month of year. What makes the plot difficult to read? How could you improve it?

```{r}

nycflights13::flights %>%
  group_by(dest, month) %>%
  count(dest, month, dep_delay) %>%  
  ggplot(mapping = aes(x = factor(month), y = dest)) +
    geom_tile(mapping = aes(fill = n)) +
    ggtitle("Departure delay by month and destination") +
    scale_fill_viridis_c() +
  labs(x = "Month", y = "Destination")


```
  
  The graph has too much information to be useful as there are too many destinations and months. It does seem like delays are influenced by destination and month (ie. December). I would explore a few graphs from example destinations, or use all destinations for one month. 

3. Why is it slightly better to use aes(x = color, y = cut) rather than aes(x = cut, y = color) in the example above?

It is usually better to put the group with the lower number of categories on the x-axis and color has less categories than cut. From a statistcal point of value, the x-axis should be the independent variable and y-axis the dependent variable.

# 7.5.2.2 Exercises

1. Instead of summarising the conditional distribution with a boxplot, you could use a frequency polygon. What do you need to consider when using cut_width() vs cut_number()? How does that impact a visualisation of the 2d distribution of carat and price?

cut_width needs the width specified, whereas cut_number needs the number within a bin specified. Either way the lines or color should be kept under 6 values or the graph gets too busy. Easier to see trends in cut_number and easier to see larger groups in cut_width


```{r}

ggplot(diamonds, aes(x = price, color = cut_width(carat, 1, boundary = 0))) + # added boundary so carat starts there
  geom_freqpoly(binwidth = 200) +
  ggtitle("geom_freqpoly with cut_width") +
  labs(color = "carat")

ggplot(diamonds, aes(price, color = cut_number(carat, 5))) + 
  geom_freqpoly(binwidth = 200) +
  ggtitle("geom_freqpoly with cut_number") +
    labs(color = "carat")

```


2. Visualise the distribution of carat, partitioned by price.

```{r}

ggplot(diamonds, aes(carat, price)) + 
  geom_bin2d() +
  ggtitle("Distribution of Carat by Price") +
  scale_fill_viridis_c()

#library(hexbin)

ggplot(diamonds, aes(carat, price)) +
  geom_hex() +
  ggtitle("Distribution of Carat by Price")+
  scale_fill_viridis_c()
  

```


3. How does the price distribution of very large diamonds compare to small diamonds? Is it as you expect, or does it surprise you?

```{r}
ggplot(diamonds, aes(carat, price)) + 
  geom_boxplot(aes(group = cut_width(carat, 0.1)))
```


The price range of diamonds gets larger as the diamonds get larger, until three carats. After 3 carats, it looks like there are too few diamonds for an accurate range.

4. Combine two of the techniques you’ve learned to visualise the combined distribution of cut, carat, and price.

```{r}

ggplot(diamonds, aes(cut_width(carat, 1, boundary = 0), price, colour = cut)) +
  geom_boxplot() +
  labs(x = "Carat") +
  scale_color_viridis_d()

```


5. Two dimensional plots reveal outliers that are not visible in one dimensional plots. For example, some points in the plot below have an unusual combination of x and y values, which makes the points outliers even though their x and y values appear normal when examined separately. Why is a scatterplot a better display than a binned plot for this case?

```{r}

ggplot(data = diamonds) +
  geom_point(mapping = aes(x = x, y = y)) +
  coord_cartesian(xlim = c(4, 11), ylim = c(4, 11)) +
  ggtitle("Scatter Plot Outliers")

ggplot(diamonds, aes(x, y)) +
  geom_bin2d(binwidth = 0.05) +
  coord_cartesian(xlim = c(4, 11), ylim = c(4, 11)) +
  ggtitle("2 Dimension Plot Outliers")

```


It looks like data with a linear relationship is better displayed with geom_point, whereas more two dimensional graphs show non- linear trends better.

