---
title: "Diamond sizes Modified"
date: 2020-06-10
output: html_document
---

```{r setup, include = FALSE}
library(ggplot2)
library(dplyr)

smaller <- diamonds %>% 
  filter(carat <= 2.5)

# global option to hide code
knitr::opts_chunk$set(
  echo = FALSE
)
```

We have data about `r nrow(diamonds)` diamonds. Only `r nrow(diamonds) - nrow(smaller)` are larger than 2.5 carats. The distribution of the remainder is shown below:

```{r intro, echo = FALSE}
smaller %>% 
  ggplot(aes(carat)) + 
  geom_freqpoly(binwidth = 0.01)
```

This example text shows that I know haw to add text for Exercise 27.3.1.3

## Code Chunks 27.4.7.1

Add a section that explores how diamond sizes vary by cut, colour, and clarity. Assume you’re writing a report for someone who doesn’t know R, and instead of setting echo = FALSE on each chunk, set a global option. I modified the global option:
```{r eval = FALSE, echo = TRUE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r codeChunks}

smaller %>% 
  ggplot(aes(color, clarity)) + 
  geom_count() +
  facet_wrap(~cut) +
  theme_bw()


```

Individual plots are of one cut quality. The colour is on x-axis and clarity on the y-axis. Larger counts have larger circles

## Code Chunks 27.4.7.2

Download diamond-sizes.Rmd from https://github.com/hadley/r4ds/tree/master/rmarkdown. Add a section that describes the largest 20 diamonds, including a table that displays their most important attributes.

```{r twenty}

diamonds %>% 
  arrange(-carat) %>%
  slice(1:20) %>%
  dplyr::rename(Carat = carat,
                Cut = cut,
                Clarity = clarity,
                Price = price,
                Color = color) %>%
  select(Carat, Cut, Color, Clarity, Price) %>%
  
  knitr::kable( ., 
  caption = "Twenty Largest Diamonds"
)


# # another way to do it
# # make column names title case
# colnames(large20) <- stringr::str_to_title(colnames(large20))
# 
# knitr::kable(
#   large20, 
#   caption = "Twenty Largest Diamonds"
# )

```

## Code Chunks 27.4.7.3

Modify diamonds-sizes.Rmd to use comma() to produce nicely formatted output. Also include the percentage of diamonds that are larger than 2.5 carats.
```{r comma}

comma <- function(x) format(x, digits = 2, big.mark = ",")

big <- diamonds %>%
  filter(carat > 2.5) %>%
  count() %>%
  pull(n)

perBig <- big/nrow(diamonds) * 100


```

Percentage of diamonds >2.5 is `r comma(perBig)` %.

## Code Chunks 27.4.7.4
Set up a network of chunks where d depends on c and b, and both b and c depend on a. Have each chunk print lubridate::now(), set cache = TRUE, then verify your understanding of caching.
```{r a, cache = TRUE}
lubridate::now()

```

```{r b, dependson= "a", cache = FALSE}

lubridate::now()

```

```{r c, dependson="a", cache = TRUE}
lubridate::now()

```

```{r d, dependson=c("c", "b")}
lubridate::now()

```

Any ideas how to get knitr::clean_cache to work?

```{r cleanCache, echo = TRUE}

# this list all caches for this document
knitr::clean_cache(FALSE)

# this deletes the caches listed in the file path
knitr::clean_cache(clean = TRUE, path = here::here("Rmarkdown", "DiamondSizes_cache"))
```

